---
title: "HiC_meta_comparisons"
author: "Ittai Eres"
date: "2023-10-04"
output: html_document
---

```{r setup, eval=FALSE, include=FALSE}
if (!requireNamespace("BiocManager", quietly=TRUE))
    install.packages("BiocManager")
BiocManager::install("HiCcompare")
BiocManager::install("TADCompare")
BiocManager::install("SpectralTAD")
BiocManager::install("multiHiCcompare")
```

```{r Library Load}
library(dplyr)
library(data.table)
library(HiCcompare)
library(multiHiCcompare)
library(SpectralTAD)
library(TADCompare)
library(BiocParallel)
library(ggplot2)
library(Hmisc)
library(stringr)
library(pheatmap)
numCores <- 96
register(MulticoreParam(workers=numCores), default=TRUE)
data("hg38_cyto")
```

## Merge significant aggregated loops from both methods and within each ancestry:
```{r}
setwd("<path>/FULL_SAMPLES/loops/")
EUR.HIC <- fread("EUR_aggregated_10kb_HICCUPS_loops.bedpe") %>% mutate(ID=paste(chrom1, start1, start2, sep=":")) %>% dplyr::select(chrom1, start1, end1, chrom2, start2, end2, ID, count) #27,184 loops
AFR.HIC <- fread("AFR_aggregated_10kb_HICCUPS_loops.bedpe") %>% mutate(ID=paste(chrom1, start1, start2, sep=":")) %>% dplyr::select(chrom1, start1, end1, chrom2, start2, end2, ID, count) #31,248 loops
EUR.FIT <- fread("EUR_aggregated_10kb_FitHiC_loops.bedpe") %>% dplyr::select(chr1, start1, end1, chr2, start2, end2, contactCount, name) #Already has a "name" column that can be used for ID #196,409 loops
AFR.FIT <- fread("AFR_aggregated_10kb_FitHiC_loops.bedpe") %>% dplyr::select(chr1, start1, end1, chr2, start2, end2, contactCount, name) #Already has a "name" column that can be used for ID #219,437 loops

EUR.HIC <- EUR.HIC %>% group_by(ID) %>% summarise(chrom1=unique(chrom1), start1=unique(start1), end1=unique(end1), chrom2=unique(chrom2), start2=unique(start2), end2=unique(end2), count=sum(count, na.rm=TRUE)) %>% dplyr::select(chrom1, start1, end1, chrom2, start2, end2, ID, count) #Reduce to 12,633 loops (>50% reduction)

AFR.HIC <- AFR.HIC %>% group_by(ID) %>% summarise(chrom1=unique(chrom1), start1=unique(start1), end1=unique(end1), chrom2=unique(chrom2), start2=unique(start2), end2=unique(end2), count=sum(count, na.rm=TRUE)) %>% dplyr::select(chrom1, start1, end1, chrom2, start2, end2, ID, count) #Reduce to 14,940 loops (>50% reduction)

EUR.FIT <- EUR.FIT %>% group_by(name) %>% summarise(chr1=unique(chr1), start1=unique(start1), end1=unique(end1), chr2=unique(chr2), start2=unique(start2), end2=unique(end2), contactCount=sum(contactCount, na.rm=TRUE)) %>% dplyr::select(chr1, start1, end1, chr2, start2, end2, name, contactCount) #Reduce to 92,848 loops (>50% reduction)

AFR.FIT <- AFR.FIT %>% group_by(name) %>% summarise(chr1=unique(chr1), start1=unique(start1), end1=unique(end1), chr2=unique(chr2), start2=unique(start2), end2=unique(end2), contactCount=sum(contactCount, na.rm=TRUE)) %>% dplyr::select(chr1, start1, end1, chr2, start2, end2, name, contactCount) #Reduce to 97,204 loops (>50% reduction)

fwrite(EUR.HIC, "EUR_aggregated_merged_10kb_HICCUPS_loops.bedpe", quote=F, col.names=T, row.names=F, sep="\t")
fwrite(AFR.HIC, "AFR_aggregated_merged_10kb_HICCUPS_loops.bedpe", quote=F, col.names=T, row.names=F, sep="\t")
fwrite(EUR.FIT, "EUR_aggregated_merged_10kb_FitHiC_loops.bedpe", quote=F, col.names=T, row.names=F, sep="\t")
fwrite(AFR.FIT, "AFR_aggregated_merged_10kb_FitHiC_loops.bedpe", quote=F, col.names=T, row.names=F, sep="\t")

#Make a version for each modality that merges all loops across ancestries!
EUR.HIC <- fread("EUR_aggregated_merged_10kb_HICCUPS_loops.bedpe")
AFR.HIC <- fread("AFR_aggregated_merged_10kb_HICCUPS_loops.bedpe")
EUR.FIT <- fread("EUR_aggregated_merged_10kb_FitHiC_loops.bedpe")
AFR.FIT <- fread("AFR_aggregated_merged_10kb_FitHiC_loops.bedpe")

#Coerce columns names & chr notations to be the same for the sake of comaprison:
colnames(EUR.FIT) <- colnames(AFR.FIT) <- colnames(EUR.HIC)
EUR.FIT$chrom1 <- gsub("chr", "", EUR.FIT$chrom1)
EUR.FIT$chrom2 <- gsub("chr", "", EUR.FIT$chrom2)
AFR.FIT$chrom1 <- gsub("chr", "", AFR.FIT$chrom1)
AFR.FIT$chrom2 <- gsub("chr", "", AFR.FIT$chrom2)
#Reconstruct IDs for FitHiC loops, as they use the midpoint of the bin instead of the bin actual coordinates:
EUR.FIT$ID <- paste(EUR.FIT$chrom1, EUR.FIT$start1, EUR.FIT$start2, sep=":")
AFR.FIT$ID <- paste(AFR.FIT$chrom1, AFR.FIT$start1, AFR.FIT$start2, sep=":")

ALL.HIC <- rbind(EUR.HIC, AFR.HIC) %>% group_by(ID) %>% summarise(chrom1=unique(chrom1), start1=unique(start1), end1=unique(end1), chrom2=unique(chrom2), start2=unique(start2), end2=unique(end2), count=sum(count, na.rm=TRUE)) %>% dplyr::select(chrom1, start1, end1, chrom2, start2, end2, ID, count) #19,970 loops--~8k less than pure aggregation (nrow(EUR.HIC) + nrow(AFR.HIC)=27,573)

ALL.FIT <- rbind(EUR.FIT, AFR.FIT) %>% group_by(ID) %>% summarise(chr1=unique(chrom1), start1=unique(start1), end1=unique(end1), chr2=unique(chrom2), start2=unique(start2), end2=unique(end2), count=sum(count, na.rm=TRUE)) %>% dplyr::select(chr1, start1, end1, chr2, start2, end2, ID, count) #138,806 loops--~50k less than pure aggregation (nrow(EUR.FIT)+nrow(AFR.FIT)=190,052)

fwrite(ALL.HIC, "ALL_aggregated_merged_10kb_HICCUPS_loops.bedpe", quote=F, col.names=T, row.names=F, sep="\t")
fwrite(ALL.FIT, "ALL_aggregated_merged_10kb_FitHiC_loops.bedpe", quote=F, col.names=T, row.names=F, sep="\t")

#Finally, write-out full loop files that include FitHiC loops and HiCCUPS loops for each ancestry, and combined:
ALL.EUR <- rbind(EUR.HIC, EUR.FIT) %>% group_by(ID) %>% summarise(chrom1=unique(chrom1), start1=unique(start1), end1=unique(end1), chrom2=unique(chrom2), start2=unique(start2), end2=unique(end2), count=sum(count, na.rm=TRUE)) %>% dplyr::select(chrom1, start1, end1, chrom2, start2, end2, ID, count) #94,200 loops--~9k less than pure aggregation (nrow(EUR.HIC)+nrow(EUR.FIT)=105,481 loops)
ALL.AFR <- rbind(AFR.HIC, AFR.FIT) %>% group_by(ID) %>% summarise(chrom1=unique(chrom1), start1=unique(start1), end1=unique(end1), chrom2=unique(chrom2), start2=unique(start2), end2=unique(end2), count=sum(count, na.rm=TRUE)) %>% dplyr::select(chrom1, start1, end1, chrom2, start2, end2, ID, count) #98,573 loops--~14k less than pure aggregation (nrow(AFR.HIC)+nrow(AFR.FIT)=112,144 loops)

ALL.ALL <- rbind(ALL.EUR, ALL.AFR) %>% group_by(ID) %>% summarise(chrom1=unique(chrom1), start1=unique(start1), end1=unique(end1), chrom2=unique(chrom2), start2=unique(start2), end2=unique(end2), count=sum(count, na.rm=TRUE)) %>% dplyr::select(chrom1, start1, end1, chrom2, start2, end2, ID, count) #140,920 loops--~52k less than pure aggregation (nrow(ALL.EUR)+nrow(ALL.AFR)=192,773 loops)

fwrite(ALL.EUR, "ALL_EUR_aggregated_merged_10kb_loops.bedpe", quote=F, col.names=T, row.names=F, sep="\t")
fwrite(ALL.AFR, "ALL_AFR_aggregated_merged_10kb_loops.bedpe", quote=F, col.names=T, row.names=F, sep="\t")
fwrite(ALL.ALL, "ALL_ALL_aggregated_merged_10kb_loops.bedpe", quote=F, col.names=T, row.names=F, sep="\t")
```

## Look at overlap of FitHiC and HiCCUPS loops across samples; note this was performed on <compute_environment>.
```{r}
setwd("<path>/FULL_SAMPLES/INTERSECTS/")

#Get all sample IDs in a vector
samples <- c(SAMPLE_01, SAMPLE_02, SAMPLE_03, SAMPLE_04, SAMPLE_05, SAMPLE_06, SAMPLE_07, SAMPLE_08, SAMPLE_09, SAMPLE_10, SAMPLE_11, SAMPLE_12, SAMPLE_13, SAMPLE_14, SAMPLE_15, SAMPLE_16, SAMPLE_17, SAMPLE_18, SAMPLE_19, SAMPLE_20, SAMPLE_21, SAMPLE_22)
#Updated version for paper that only includes included samples:
samples <- c(SAMPLE_05, SAMPLE_09, SAMPLE_10, SAMPLE_11, SAMPLE_12, SAMPLE_13, SAMPLE_14, SAMPLE_15, SAMPLE_16, SAMPLE_17, SAMPLE_18, SAMPLE_19, SAMPLE_20, SAMPLE_21, SAMPLE_22)
resolutions <- c(10, 5, 2)

overlaps <- data.frame()

for (sample in samples) {
  for (resolution in resolutions){
  tmp.df <- fread(paste0(sample, "_", resolution,"kb_loopcaller_intersects.txt"))
  colnames(tmp.df) <- c("FitHiC", "Shared", "HiCCUPS")
  tmp.df$sample <- sample
  tmp.df$resolution <- resolution*1000
  tmp.df <- melt(tmp.df, id.vars=c("sample", "resolution"))
  overlaps <- rbind(overlaps, tmp.df)
}
}

overlaps$sample <- as.factor(overlaps$sample)

#10kb plot
ggplot(data=dplyr::filter(overlaps, resolution==10000), aes(x=sample, group=sample, y=value, fill=variable)) + geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Sample ID") + ylab("Significant Loop Count") + ggtitle("Intra-Sample Comparison of Significant Loops, FitHiC vs. HiCCUPS, 10kb") + guides(fill=guide_legend(title="Loop Origin"))
ggsave("../QCFIGS/intra_method_loops_10kb_counts.pdf")
ggplot(data=dplyr::filter(overlaps, resolution==10000), aes(x=sample, group=sample, y=value, fill=variable)) + geom_bar(stat="identity", position="fill") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Sample ID") + ylab("Percentage of Significant Loops") + ggtitle("Intra-Sample Comparison of Significant Loops, FitHiC vs. HiCCUPS, 10kb") + guides(fill=guide_legend(title="Loop Origin")) + 
  scale_y_continuous(labels = scales::percent_format())
ggsave("../QCFIGS/intra_method_loops_10kb_percentages.pdf")

#5kb plot
ggplot(data=dplyr::filter(overlaps, resolution==5000), aes(x=sample, group=sample, y=value, fill=variable)) + geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Sample ID") + ylab("Significant Loop Count") + ggtitle("Intra-Sample Comparison of Significant Loops, FitHiC vs. HiCCUPS, 5kb") + guides(fill=guide_legend(title="Loop Origin"))
ggsave("../QCFIGS/intra_method_loops_5kb_counts.pdf")
ggplot(data=dplyr::filter(overlaps, resolution==5000), aes(x=sample, group=sample, y=value, fill=variable)) + geom_bar(stat="identity", position="fill") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Sample ID") + ylab("Percentage of Significant Loops") + ggtitle("Intra-Sample Comparison of Significant Loops, FitHiC vs. HiCCUPS, 5kb") + guides(fill=guide_legend(title="Loop Origin")) + 
  scale_y_continuous(labels = scales::percent_format())
ggsave("../QCFIGS/intra_method_loops_5kb_percentages.pdf")

#2kb plot
ggplot(data=dplyr::filter(overlaps, resolution==2000), aes(x=sample, group=sample, y=value, fill=variable)) + geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Sample ID") + ylab("Significant Loop Count") + ggtitle("Intra-Sample Comparison of Significant Loops, FitHiC vs. HiCCUPS, 2kb") + guides(fill=guide_legend(title="Loop Origin"))
ggsave("../QCFIGS/intra_method_loops_2kb_counts.pdf")
ggplot(data=dplyr::filter(overlaps, resolution==2000), aes(x=sample, group=sample, y=value, fill=variable)) + geom_bar(stat="identity", position="fill") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Sample ID") + ylab("Percentage of Significant Loops") + ggtitle("Intra-Sample Comparison of Significant Loops, FitHiC vs. HiCCUPS, 2kb") + guides(fill=guide_legend(title="Loop Origin")) + 
  scale_y_continuous(labels = scales::percent_format())
ggsave("../QCFIGS/intra_method_loops_2kb_percentages.pdf")

#Check for merged aggregated loops across all samples, stratified by ancestry as well as combined:
fit.overlaps <- fread("AFR_vs_EUR_FitHiC_intersects.txt") %>% mutate(African=Unique_to_1, European=Unique_to_2) %>% select(African, Shared, European) %>% mutate(Method="FitHiC")
hic.overlaps <- fread("AFR_vs_EUR_HICCUPS_intersects.txt") %>% mutate(African=Unique_to_1, European=Unique_to_2) %>% select(African, Shared, European) %>% mutate(Method="HiCCUPS")
anc.overlaps.all <- fread("AFR_vs_EUR_all_intersects_10kb.txt") %>% mutate(African=Unique_to_1, European=Unique_to_2) %>% select(African, Shared, European) %>% mutate(Method="Both Merged")

all.overlaps <- rbind(fit.overlaps, hic.overlaps, anc.overlaps.all)
all.overlaps <- melt(all.overlaps, id.vars = "Method")
all.overlaps$Method <- factor(all.overlaps$Method, levels=c("HiCCUPS", "FitHiC", "Both Merged"))

ggplot(data=all.overlaps, aes(x=Method, group=Method, y=value, fill=variable)) + geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Loop-Calling Method") + ylab("Significant Loop Count") + ggtitle("Intra-Ancestry Comparison of Merged Significant Loops, FitHiC vs. HiCCUPS, 10kb") + guides(fill=guide_legend(title="Loop Origin"))
ggsave("../QCFIGS/intra_ancestry_loops_10kb_counts.pdf")
ggplot(data=all.overlaps, aes(x=Method, group=Method, y=value, fill=variable)) + geom_bar(stat="identity", position="fill") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Loop-Calling Method") + ylab("Percentage of Significant Loops") + ggtitle("Intra-Ancestry Comparison of Merged Significant Loops, FitHiC vs. HiCCUPS, 10kb") + guides(fill=guide_legend(title="Loop Origin")) + 
  scale_y_continuous(labels = scales::percent_format())
ggsave("../QCFIGS/intra_ancestry_loops_10kb_percentages.pdf")


method.overlaps.all <- fread("FitHiC_vs_HICCUPS_ALL_10kb.txt") %>% mutate(FitHiC=Unique_to_1, HiCCUPS=Unique_to_2) %>% select(FitHiC, Shared, HiCCUPS) %>% mutate(sample="ALL") %>% melt(., id.vars="sample")
overlaps10kb <- filter(overlaps, resolution==10000) %>% select(-resolution)
overlaps10kb <- rbind(overlaps10kb, method.overlaps.all)

ggplot(data=overlaps10kb, aes(x=sample, group=sample, y=value, fill=variable)) + geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Sample ID") + ylab("Significant Loop Count") + ggtitle("Intra-Sample Comparison of Significant Loops, FitHiC vs. HiCCUPS, 10kb") + guides(fill=guide_legend(title="Loop Origin"))
ggsave("../QCFIGS/intra_method_merged_loops_10kb_counts.pdf")
ggplot(data=overlaps10kb, aes(x=sample, group=sample, y=value, fill=variable)) + geom_bar(stat="identity", position="fill") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Sample ID") + ylab("Percentage of Significant Loops") + ggtitle("Intra-Sample Comparison of Significant Loops, FitHiC vs. HiCCUPS, 10kb") + guides(fill=guide_legend(title="Loop Origin")) + 
  scale_y_continuous(labels = scales::percent_format())
ggsave("../QCFIGS/intra_method_merged_loops_10kb_percentages.pdf")
```

### Look at contact distance distributions per sample, methodology, and resolution. Again, performed on <compute_environment>:
```{r}
library(data.table)
library(ggplot2)
library(dplyr)

# Samples/resolutions from your earlier block
samples <- c(SAMPLE_05, SAMPLE_09, SAMPLE_10, SAMPLE_11, SAMPLE_12, SAMPLE_13, SAMPLE_14, SAMPLE_15,
             SAMPLE_16, SAMPLE_17, SAMPLE_18, SAMPLE_19, SAMPLE_20, SAMPLE_21, SAMPLE_22)
resolutions <- c(10000, 5000, 2000)

# Toggle display cap without removing data
cap_y_at_2e6 <- TRUE
y_cap <- 2e6

# Helper: read BEDPE-like loop file and compute contact distance
get_contact_distances <- function(file, sample_id, resolution, method, header = TRUE) {
  dt <- fread(file, header = header, sep = "\t", quote = "", data.table = FALSE)
  
  if (ncol(dt) < 6) {
    stop(paste("File has fewer than 6 columns:", file))
  }
  
  dt <- dt[, 1:6]
  colnames(dt) <- c("chr1", "start1", "end1", "chr2", "start2", "end2")
  
  dt$chr1 <- gsub("^chr", "", dt$chr1)
  dt$chr2 <- gsub("^chr", "", dt$chr2)
  
  dt$start1 <- as.numeric(dt$start1)
  dt$end1   <- as.numeric(dt$end1)
  dt$start2 <- as.numeric(dt$start2)
  dt$end2   <- as.numeric(dt$end2)
  
  dt <- dt %>%
    filter(!is.na(start1), !is.na(end1), !is.na(start2), !is.na(end2)) %>%
    filter(chr1 == chr2)
  
  dt$mid1 <- (dt$start1 + dt$end1) / 2
  dt$mid2 <- (dt$start2 + dt$end2) / 2
  dt$contact_distance_bp <- abs(dt$mid2 - dt$mid1)
  
  dt$sample <- as.factor(sample_id)
  dt$resolution <- resolution
  dt$method <- method
  
  dt[, c("sample", "resolution", "method", "chr1", "chr2", "contact_distance_bp")]
}

# Build long table
loop_dists <- data.frame()

for (sample in samples) {
  for (resolution in resolutions) {
    res_kb <- resolution / 1000
    
    fithic_file <- paste0(
      "<path>/FULL_SAMPLES/FITHIC_2026/LOOPS/BEDPES/",
      sample, ".fithic.", resolution, ".loops.bedpe"
    )
    
    hiccups_file <- paste0(
      "<path>/FULL_SAMPLES/loops/",
      "loops_", res_kb, "kb", sample, ".bedpe"
    )
    
    tmp_fithic <- get_contact_distances(
      file = fithic_file,
      sample_id = sample,
      resolution = resolution,
      method = "FitHiC",
      header = TRUE
    )
    
    tmp_hiccups <- get_contact_distances(
      file = hiccups_file,
      sample_id = sample,
      resolution = resolution,
      method = "HiCCUPS",
      header = TRUE
    )
    
    loop_dists <- rbind(loop_dists, tmp_fithic, tmp_hiccups)
  }
}

loop_dists$sample <- factor(loop_dists$sample, levels = as.character(samples))

# Plotting function: one plot per method per resolution
make_violin_plot <- function(df, method_name, resolution_value, cap_y = FALSE, ymax = 2e6) {
  p <- ggplot(
    data = df %>% filter(method == method_name, resolution == resolution_value),
    aes(x = sample, y = contact_distance_bp, group = sample)
  ) +
    geom_violin(fill = "lightblue", color = "black", trim = TRUE, scale = "width") +
    geom_boxplot(
      width = 0.12,
      fill = "white",
      color = "black",
      outlier.shape = NA
    ) +
    scale_y_continuous(
      breaks = c(0, 5e5, 1e6, 1.5e6, 2e6),
      labels = c(
        "0",
        expression(5 %*% 10^5),
        expression(1 %*% 10^6),
        expression(1.5 %*% 10^6),
        expression(2 %*% 10^6)
      )
    ) +
    theme_bw() +
    theme(
      axis.text.x = element_text(angle = 90, hjust = 1),
      plot.title = element_text(face = "bold")
    ) +
    xlab("Sample ID") +
    ylab("Contact Distance (bp)") +
    ggtitle(paste0(method_name, " Loop Contact Distance Distributions, ", resolution_value / 1000, " kb"))
  
  if (cap_y) {
    p <- p + coord_cartesian(ylim = c(0, ymax))
  }
  
  p
}

# Generate separate plots per resolution for FitHiC
for (res in resolutions) {
  p <- make_violin_plot(
    df = loop_dists,
    method_name = "FitHiC",
    resolution_value = res,
    cap_y = cap_y_at_2e6,
    ymax = y_cap
  )
  
  suffix <- ifelse(cap_y_at_2e6, "_ymax2e6", "")
  ggsave(
    filename = paste0("../QCFIGS/fithic_contact_distance_", res / 1000, "kb_violin", suffix, ".pdf"),
    plot = p,
    width = 12,
    height = 6
  )
}

# Generate separate plots per resolution for HiCCUPS
for (res in resolutions) {
  p <- make_violin_plot(
    df = loop_dists,
    method_name = "HiCCUPS",
    resolution_value = res,
    cap_y = cap_y_at_2e6,
    ymax = y_cap
  )
  
  suffix <- ifelse(cap_y_at_2e6, "_ymax2e6", "")
  ggsave(
    filename = paste0("../QCFIGS/hiccups_contact_distance_", res / 1000, "kb_violin", suffix, ".pdf"),
    plot = p,
    width = 12,
    height = 6
  )
}
```

### HiCRes results & other QC metrics for library complexity, again on <compute_environment>:
```{r}
##### Get HiCRes results as one panel
library(magick)
library(cowplot)
library(patchwork)

base_dir <- "<path>/FULL_SAMPLES/hicres_results"

dirs <- list.dirs(base_dir, recursive = FALSE)
dirs <- dirs[!grepl("old|QC_|SAMPLE_01|SAMPLE_02|SAMPLE_03|SAMPLE_04|4200", dirs)]

get_sample <- function(path) {
  sub("_HiCRes_outputs$", "", basename(path))
}

plots <- list()

for (d in dirs) {
  img_path <- file.path(d, "prediction.png")
  if (!file.exists(img_path)) next
  
  sample <- get_sample(d)
  
  img <- image_read(img_path)
  img <- image_trim(img)
  
  p <- ggdraw() +
    draw_label(
      sample,
      x = 0.5, y = 0.95,
      hjust = 0.5, vjust = 1,
      fontface = "bold",
      size = 12
    ) +
    draw_image(
      img,
      x = 0, y = 0,
      width = 1, height = 0.98,
      hjust = 0, vjust = 0
    )
  
  plots[[sample]] <- p
}

combined <- wrap_plots(plots, ncol = 4)

ggsave(
  filename = "../QCFIGS/hicres_combined_panels.pdf",
  plot = combined,
  width = 16,
  height = 20
)

##### Attempt to make it in a single figure:
library(data.table)
library(ggplot2)
library(dplyr)
library(stringr)
library(scales)

all_preds <- data.frame()

for (d in dirs) {
  pred_file <- file.path(d, "prediction.txt")
  if (!file.exists(pred_file)) next
  
  sample <- get_sample(d)
  tmp <- fread(pred_file)
  
  # enforce names explicitly
  colnames(tmp) <- c("Valid_Reads", "Predicted_Resolution")
  
  tmp$sample <- sample
  tmp$Valid_Reads_M <- tmp$Valid_Reads / 1e6
  
  all_preds <- rbind(all_preds, tmp)
}

all_preds$sample <- factor(all_preds$sample, levels = sort(unique(all_preds$sample)))

p_overlay <- ggplot(
  all_preds,
  aes(x = Valid_Reads_M, y = Predicted_Resolution, color = sample, group = sample)
) +
  geom_line(linewidth = 0.9, alpha = 0.9) +
  theme_bw() +
  xlab("Valid read pairs (M)") +
  ylab("Predicted resolution (bp)") +
  ggtitle("HiCRes Predicted Resolution Curves Across Samples") +
  scale_y_continuous(labels = comma_format()) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  ) + coord_cartesian(ylim=c(0, 500000))

ggsave(
  "../QCFIGS/hicres_overlay_all_samples_from_prediction_txt.pdf",
  plot = p_overlay,
  width = 12,
  height = 8
)

p_overlay_log <- ggplot(
  all_preds,
  aes(x = Valid_Reads_M, y = Predicted_Resolution, color = sample, group = sample)
) +
  geom_line(linewidth = 0.9, alpha = 0.9) +
  theme_bw() +
  xlab("Valid read pairs (M)") +
  ylab("Predicted resolution (bp, log scale)") +
  ggtitle("HiCRes Predicted Resolution Curves Across Samples") +
  scale_y_log10(
    breaks = c(2e3, 5e3, 1e4, 2e4, 1e5, 5e5),
    labels = comma_format()
  ) +
  theme(
    plot.title = element_text(face = "bold"),
    legend.position = "right"
  ) +
  coord_cartesian(ylim = c(1e3, 5e5))
ggsave(
  "../QCFIGS/hicres_overlay_all_samples_from_prediction_txt_logY.pdf",
  plot = p_overlay_log,
  width = 12,
  height = 8
)


##### Take a look at unique valid pairs and inferred resolution:
qc_base_dir <- file.path(base_dir, "QC_outputs")
# -------------------------
# Helper functions
# -------------------------

read_valid_pairs <- function(sample_id) {
  qc_dir <- file.path(qc_base_dir, paste0(sample_id, "_QC_outputs"))
  qc_file <- list.files(qc_dir, pattern = "^hicup_filter_summary_.*\\.txt$", full.names = TRUE)
  
  if (length(qc_file) != 1) {
    stop(paste("Expected exactly 1 HiCUP summary file for sample", sample_id,
               "but found", length(qc_file)))
  }
  
  dt <- fread(qc_file)
  if (!"Valid_pairs" %in% colnames(dt)) {
    stop(paste("Valid_pairs column not found in", qc_file))
  }
  
  data.frame(
    sample = as.character(sample_id),
    Valid_pairs = as.numeric(dt$Valid_pairs[1])
  )
}

read_prediction <- function(sample_id) {
  pred_file <- file.path(
    base_dir,
    paste0(sample_id, "_HiCRes_outputs"),
    "prediction.txt"
  )
  
  if (!file.exists(pred_file)) {
    stop(paste("Missing prediction.txt for sample", sample_id))
  }
  
  dt <- fread(pred_file)
  colnames(dt) <- c("Valid_Reads", "Predicted_Resolution")
  dt$sample <- as.character(sample_id)
  dt
}

# nearest predicted resolution for actual valid-pair count
get_resolution_at_sample_depth <- function(pred_dt, valid_pairs) {
  pred_dt <- pred_dt %>%
    mutate(diff = abs(Valid_Reads - valid_pairs)) %>%
    arrange(diff)
  
  pred_dt$Predicted_Resolution[1]
}

# optional linear interpolation if actual valid_pairs falls between grid points
get_resolution_interpolated <- function(pred_dt, valid_pairs) {
  approx(
    x = pred_dt$Valid_Reads,
    y = pred_dt$Predicted_Resolution,
    xout = valid_pairs,
    rule = 2
  )$y
}

# -------------------------
# Build summary tables
# -------------------------

valid_pairs_df <- do.call(
  rbind,
  lapply(samples, read_valid_pairs)
)

prediction_list <- lapply(samples, read_prediction)
all_preds <- do.call(rbind, prediction_list)

resolution_df <- lapply(samples, function(s) {
  s_char <- as.character(s)
  
  vp <- valid_pairs_df %>%
    filter(sample == s_char) %>%
    pull(Valid_pairs)
  
  pred_dt <- all_preds %>%
    filter(sample == s_char)
  
  # choose one:
  # predicted_res <- get_resolution_at_sample_depth(pred_dt, vp)     # nearest point
  predicted_res <- get_resolution_interpolated(pred_dt, vp)          # smoother / better
  
  data.frame(
    sample = s_char,
    Valid_pairs = vp,
    Predicted_Resolution = predicted_res
  )
}) %>% bind_rows()

# preserve sample order
valid_pairs_df$sample <- factor(valid_pairs_df$sample, levels = as.character(samples))
resolution_df$sample <- factor(resolution_df$sample, levels = as.character(samples))

# -------------------------
# Plot 1: valid pairs barplot
# -------------------------

p_valid_pairs <- ggplot(valid_pairs_df, aes(x = sample, y = Valid_pairs)) +
  geom_col(fill = "steelblue") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    plot.title = element_text(face = "bold")
  ) +
  xlab("Sample ID") +
  ylab("Unique valid pairs") +
  ggtitle("Unique Valid Pairs per Sample") +
  scale_y_continuous(labels = comma_format(), breaks=c(250000000, 500000000, 750000000, 1000000000, 1250000000))

ggsave(
  "../QCFIGS/hicup_valid_pairs_per_sample.pdf",
  plot = p_valid_pairs,
  width = 10,
  height = 6
)

# -------------------------
# Plot 2: inferred HiCRes resolution barplot
# -------------------------

p_resolution <- ggplot(resolution_df, aes(x = sample, y = Predicted_Resolution)) +
  geom_col(fill = "firebrick2") +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    plot.title = element_text(face = "bold")
  ) +
  xlab("Sample ID") +
  ylab("Predicted resolution (bp)") +
  ggtitle("HiCRes Predicted Resolution at Observed Valid-Pair Depth") +
  scale_y_continuous(labels = comma_format(), breaks=c(1e3, 2e3, 3e3, 4e3, 5e3, 6e3, 7e3, 8e3, 9e3, 10e3)) + coord_cartesian(ylim=c(0, 10001))

ggsave(
  "../QCFIGS/hicres_predicted_resolution_at_observed_depth.pdf",
  plot = p_resolution,
  width = 10,
  height = 6
)

# -------------------------
# Optional: merged summary table
# -------------------------

summary_df <- valid_pairs_df %>%
  left_join(resolution_df %>% select(sample, Predicted_Resolution), by = "sample")

write.table(
  summary_df,
  file = "../QCFIGS/hicres_valid_pairs_and_resolution_summary.tsv",
  sep = "\t",
  row.names = FALSE,
  quote = FALSE
)
```

### HiC-Pro contact pairing statistics, again on <compute_environment>
```{r}
###### HiC-Pro Contact Statistics BarPlots
library(data.table)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)

base_dir <- "<path>/FULL_SAMPLES"

sample_order <- samples

contact_categories <- c(
  "Unique: cis <= 20Kbp",
  "Unique: cis > 20Kbp",
  "Unique: trans",
  "Duplicate read pairs"
)

contact_colors <- c(
  "Unique: cis <= 20Kbp" = "#143fd6",  # dark blue
  "Unique: cis > 20Kbp" = "#7893e2",   # light blue
  "Unique: trans" = "#06a02d",         # green
  "Duplicate read pairs" = "#ada7a7"   # gray
)

read_contact_file <- function(sample_id, base_dir) {
  infile <- file.path(
    base_dir,
    as.character(sample_id),
    "multiqc",
    "mqc_hicpro_contact_plot_1.txt"
  )
  
  if (!file.exists(infile)) {
    warning(paste("Missing file for sample", sample_id, ":", infile))
    return(NULL)
  }
  
  dt <- fread(infile)
  
  required_cols <- c(
    "Sample",
    "Unique: cis <= 20Kbp",
    "Unique: cis > 20Kbp",
    "Unique: trans",
    "Duplicate read pairs"
  )
  
  if (!all(required_cols %in% colnames(dt))) {
    stop(
      paste(
        "File", infile, "is missing required columns. Found:",
        paste(colnames(dt), collapse = ", ")
      )
    )
  }
  
  # Expect one row per sample file; if multiple rows exist, sum them
  out <- data.frame(
    sample = as.character(sample_id),
    `Unique: cis <= 20Kbp` = sum(as.numeric(dt[["Unique: cis <= 20Kbp"]]), na.rm = TRUE),
    `Unique: cis > 20Kbp`  = sum(as.numeric(dt[["Unique: cis > 20Kbp"]]), na.rm = TRUE),
    `Unique: trans`        = sum(as.numeric(dt[["Unique: trans"]]), na.rm = TRUE),
    `Duplicate read pairs` = sum(as.numeric(dt[["Duplicate read pairs"]]), na.rm = TRUE),
    stringsAsFactors = FALSE
  )
  
  out
}

# Read all sample files
contact_df_wide <- bind_rows(
  lapply(sample_order, read_contact_file, base_dir = base_dir)
)

# Preserve order
contact_df_wide$sample <- factor(contact_df_wide$sample, levels = as.character(sample_order))

# Fix colnames
colnames(contact_df_wide)[2:5] <- c("Unique: cis <= 20Kbp", "Unique: cis > 20Kbp", "Unique: trans", "Duplicate read pairs")

# Write summary table
write.table(
  contact_df_wide,
  file = "../QCFIGS/hicpro_contact_summary.tsv",
  sep = "\t",
  row.names = FALSE,
  quote = FALSE
)

# Convert to long format
contact_df_long <- contact_df_wide %>%
  pivot_longer(
    cols = all_of(contact_categories),
    names_to = "category",
    values_to = "count"
  )

contact_df_long$category <- factor(contact_df_long$category, levels = contact_categories)

# Add percentages
contact_df_long <- contact_df_long %>%
  group_by(sample) %>%
  mutate(
    total = sum(count, na.rm = TRUE),
    percent = 100 * count / total
  ) %>%
  ungroup()

contact_df_long$category <- factor(
  contact_df_long$category,
  levels = rev(contact_categories)
)

# -------------------------
# Plot 1: stacked counts
# -------------------------
p_counts <- ggplot(contact_df_long, aes(x = sample, y = count, fill = category)) +
  geom_col() +
  scale_fill_manual(values = contact_colors) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    plot.title = element_text(face = "bold")
  ) +
  xlab("Sample ID") +
  ylab("Read pair count") +
  ggtitle("HiC-Pro Contact Statistics per Sample") +
  scale_y_continuous(labels = comma_format()) +
  guides(fill = guide_legend(title = NULL))

ggsave(
  "../QCFIGS/hicpro_contact_statistics_counts_stacked.pdf",
  plot = p_counts,
  width = 12,
  height = 6
)

# -------------------------
# Plot 2: stacked percentages
# -------------------------
p_percent <- ggplot(contact_df_long, aes(x = sample, y = percent, fill = category)) +
  geom_col() +
  scale_fill_manual(values = contact_colors) +
  theme_bw() +
  theme(
    axis.text.x = element_text(angle = 90, hjust = 1),
    plot.title = element_text(face = "bold")
  ) +
  xlab("Sample ID") +
  ylab("Percentage of read pairs") +
  ggtitle("HiC-Pro Contact Statistics per Sample (%)") +
  scale_y_continuous(labels = function(x) paste0(x, "%")) +
  guides(fill = guide_legend(title = NULL))

ggsave(
  "../QCFIGS/hicpro_contact_statistics_percent_stacked.pdf",
  plot = p_percent,
  width = 12,
  height = 6
)
```


### Stitch together saddleplots across all samples. Do this on <compute_environment> since I had to install pdfjam:
```{unix}
cd <path>/FULL_SAMPLES/QCFIGS

  pdfjam [0-9]*.1000000_saddle_plot.pdf \
  --nup 5x3 \
  --landscape \
  --paper a1paper \
  --outfile all_samples_saddle_plot_panel.pdf
```


### Finally, look at loop overlaps vs HiChIP and E2G loops!:
```{r}
setwd("<path>/FULL_SAMPLES/loops/MultiomeHiC_revisions/E2G")

#First, re-format these files to be BEDPE-style
files <- c("Endothelium.EnhancerPredictionsThresholded.tsv", "Fibroblast.EnhancerPredictionsThresholded.tsv", "Myeloid.EnhancerPredictionsThresholded.tsv", "VSMC_encode_e2g_predictions_threshold.171.tsv")
for (file in files) {
  newname <- paste0(gsub(".tsv", "", file), ".bedpe")
  tmp.file <- fread(file) %>% dplyr::select(chr, start, end, TargetGeneTSS) %>% mutate(chr1=chr, start1=start, end1=end, chr2=chr, start2=TargetGeneTSS, end2=TargetGeneTSS+10000) %>% dplyr::select(chr1, start1, end1, chr2, start2, end2)
  fwrite(tmp.file, newname, quote=F, col.names=T, row.names=F, sep="\t")
}

#Files in the HiChIP_bedpe folder are already formatted properly! Both have headers!
#<path>/FULL_SAMPLES/loops/MultiomeHiC_revisions/HiChIP_bedpe

#Read-in and format summarization files:
setwd("<path>/FULL_SAMPLES/INTERSECTS/")
overlaps <- data.frame()
files <- c("Endothelium.EnhancerPredictionsThresholded.bedpe", "Fibroblast.EnhancerPredictionsThresholded.bedpe", "Myeloid.EnhancerPredictionsThresholded.bedpe", "VSMC_encode_e2g_predictions_threshold.171.bedpe")
files2 <- c("FITHiChIP_EC_5kb.interactions_FitHiC_Q0.01_MergeNearContacts_filteredwDistance.bed", "FITHiChIP_SMC_5kb.interactions_FitHiC_Q0.01_MergeNearContacts_filteredwDistance.bed")

for (file in files) {
  name <- gsub("\\..*", "", file)
  tmp.df <- fread(paste0("ALL_vs_E2G_", file,"_intersects.txt"))
  colnames(tmp.df) <- c("HiC", "Shared", "E2G")
  tmp.df$dataset <- paste0(name, "_E2G")
  tmp.df <- melt(tmp.df, id.vars=c("dataset"))
  overlaps <- rbind(overlaps, tmp.df)
}

overlaps$dataset <- gsub("_.*", "", overlaps$dataset)

for (file in files2) {
  name <- gsub("\\..*", "", file)
  name <- gsub("(.*)_(.*)_(.*)", "\\2", name)
  tmp.df <- fread(paste0("ALL_vs_HiChIP_", file,"_intersects.txt"))
  colnames(tmp.df) <- c("HiC", "Shared", "HiChIP")
  tmp.df$dataset <- name
  tmp.df <- melt(tmp.df, id.vars=c("dataset"))
  overlaps <- rbind(overlaps, tmp.df)
}

overlaps$dataset <- factor(overlaps$dataset, levels=c("Endothelium", "Fibroblast", "Myeloid", "VSMC", "EC", "SMC"))

#Plot these!
ggplot(data=overlaps, aes(x=dataset, group=dataset, y=value, fill=variable)) + geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Cell Type") + ylab("Significant Loop Count") + ggtitle("Comparison of Significant Loops, Hi-C vs. E2G & HiChIP") + guides(fill=guide_legend(title="Loop Origin"))
ggsave("../QCFIGS/intra_assay_loops_counts.pdf")
ggplot(data=overlaps, aes(x=dataset, group=dataset, y=value, fill=variable)) + geom_bar(stat="identity", position="fill") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Cell Type") + ylab("Percentage of Significant Loops") + ggtitle("Comparison of Significant Loops, Hi-C vs. E2G & HiChIP") + guides(fill=guide_legend(title="Loop Origin")) + 
  scale_y_continuous(labels = scales::percent_format())
ggsave("../QCFIGS/intra_assay_loops_percentages.pdf")

## See if/how this is different for the SHARED loops across FitHiC & HiCCUPS:
overlaps <- data.frame()
files <- c("Endothelium.EnhancerPredictionsThresholded.bedpe", "Fibroblast.EnhancerPredictionsThresholded.bedpe", "Myeloid.EnhancerPredictionsThresholded.bedpe", "VSMC_encode_e2g_predictions_threshold.171.bedpe")
files2 <- c("FITHiChIP_EC_5kb.interactions_FitHiC_Q0.01_MergeNearContacts_filteredwDistance.bed", "FITHiChIP_SMC_5kb.interactions_FitHiC_Q0.01_MergeNearContacts_filteredwDistance.bed")

for (file in files) {
  name <- gsub("\\..*", "", file)
  tmp.df <- fread(paste0("SHARED_ALL_vs_E2G_", file,"_intersects.txt"))
  colnames(tmp.df) <- c("HiC", "Shared", "E2G")
  tmp.df$dataset <- paste0(name, "_E2G")
  tmp.df <- melt(tmp.df, id.vars=c("dataset"))
  overlaps <- rbind(overlaps, tmp.df)
}

overlaps$dataset <- gsub("_.*", "", overlaps$dataset)

for (file in files2) {
  name <- gsub("\\..*", "", file)
  name <- gsub("(.*)_(.*)_(.*)", "\\2", name)
  tmp.df <- fread(paste0("SHARED_ALL_vs_HiChIP_", file,"_intersects.txt"))
  colnames(tmp.df) <- c("HiC", "Shared", "HiChIP")
  tmp.df$dataset <- name
  tmp.df <- melt(tmp.df, id.vars=c("dataset"))
  overlaps <- rbind(overlaps, tmp.df)
}

overlaps$dataset <- factor(overlaps$dataset, levels=c("Endothelium", "Fibroblast", "Myeloid", "VSMC", "EC", "SMC"))

#Plot these!
ggplot(data=overlaps, aes(x=dataset, group=dataset, y=value, fill=variable)) + geom_bar(stat="identity") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Cell Type") + ylab("Shared Significant Loop Count") + ggtitle("Comparison of Shared Significant Loops, Hi-C vs. E2G & HiChIP") + guides(fill=guide_legend(title="Loop Origin"))
ggsave("../QCFIGS/SHARED_intra_assay_loops_counts.pdf")
ggplot(data=overlaps, aes(x=dataset, group=dataset, y=value, fill=variable)) + geom_bar(stat="identity", position="fill") + theme(axis.text.x=element_text(angle=90, hjust=1)) + xlab("Cell Type") + ylab("Percentage of Shared Significant Loops") + ggtitle("Comparison of Shared Significant Loops, Hi-C vs. E2G & HiChIP") + guides(fill=guide_legend(title="Loop Origin")) + 
  scale_y_continuous(labels = scales::percent_format())
ggsave("../QCFIGS/SHARED_intra_assay_loops_percentages.pdf")
```

## Code for running intersections--note this is run on ieu20, not here:
```{unix}
# A loop to go through all samples and compare their results across Hiccups & Fithic, at different resolutions. This was the OLD version:
for SAMPLE in SAMPLE_01 SAMPLE_02 SAMPLE_03 SAMPLE_04 SAMPLE_05 SAMPLE_06 SAMPLE_07 SAMPLE_08 SAMPLE_09 SAMPLE_10 SAMPLE_11 SAMPLE_12 SAMPLE_13 SAMPLE_14 SAMPLE_15 SAMPLE_16 SAMPLE_17 SAMPLE_18 SAMPLE_19 SAMPLE_20 SAMPLE_21 SAMPLE_22; do
    for RES in 10000 5000 2000; do
        RES_KB=$(( RES / 1000 ))
        echo "Using resolution ${RES} bp, which is ${RES_KB} kb. Analyzing sample ${SAMPLE}."
        
        ./Venn_Interactions.r \
            --FileList=<path>/FULL_SAMPLES/${SAMPLE}/fithic/${RES}/${SAMPLE}_${RES}.bedpe,<path>/FULL_SAMPLES/loops/loops_${RES_KB}kb${SAMPLE}.bedpe \
            --HeaderList=0,1 --offset=0 \
            --Labels=${SAMPLE}_FitHiC,${SAMPLE}_HiCCUPs \
            --OutDir=<path>/FULL_SAMPLES/INTERSECTS/ \
            --Title="${SAMPLE}_${RES_KB}kb_loopcaller_intersects"
    done
done

# A loop to go through all samples and compare their results across Hiccups & Fithic, at different resolutions. UPDATED version post FitHiC loop call changes:
for SAMPLE in SAMPLE_01 SAMPLE_02 SAMPLE_03 SAMPLE_04 SAMPLE_05 SAMPLE_06 SAMPLE_07 SAMPLE_08 SAMPLE_09 SAMPLE_10 SAMPLE_11 SAMPLE_12 SAMPLE_13 SAMPLE_14 SAMPLE_15 SAMPLE_16 SAMPLE_17 SAMPLE_18 SAMPLE_19 SAMPLE_20 SAMPLE_21 SAMPLE_22; do
    for RES in 10000 5000 2000; do
        RES_KB=$(( RES / 1000 ))
        echo "Using resolution ${RES} bp, which is ${RES_KB} kb. Analyzing sample ${SAMPLE}."
        
        ./Venn_Interactions.r \
            --FileList=<path>/FULL_SAMPLES/FITHIC_2026/LOOPS/BEDPES/${SAMPLE}.fithic.${RES}.loops.bedpe,<path>/FULL_SAMPLES/loops/loops_${RES_KB}kb${SAMPLE}.bedpe \
            --HeaderList=1,1 --offset=0 \
            --Labels=${SAMPLE}_FitHiC,${SAMPLE}_HiCCUPs \
            --OutDir=<path>/FULL_SAMPLES/INTERSECTS/ \
            --Title="${SAMPLE}_${RES_KB}kb_loopcaller_intersects"
    done
done


#Rest of comparisons are 10kb resolution only
RES=10000
RES_KB=10

#Compare across HiCCUPS for AFR vs. EUR
./Venn_Interactions.r --FileList=<path>/FULL_SAMPLES/loops/AFR_aggregated_merged_10kb_HICCUPS_loops.bedpe,<path>/FULL_SAMPLES/loops/EUR_aggregated_merged_10kb_HICCUPS_loops.bedpe --HeaderList=1,1 --Labels=AFR_HICCUPS,EUR_HICCUPS --OutDir=<path>/FULL_SAMPLES/INTERSECTS/ --Title="AFR_vs_EUR_HICCUPS_intersects"

#Compare across FitHiC for AFR vs. EUR
./Venn_Interactions.r --FileList=<path>/FULL_SAMPLES/loops/AFR_aggregated_merged_10kb_FitHiC_loops.bedpe,<path>/FULL_SAMPLES/loops/EUR_aggregated_merged_10kb_FitHiC_loops.bedpe --HeaderList=1,1 --Labels=AFR_FitHiC,EUR_FitHiC --OutDir=<path>/FULL_SAMPLES/INTERSECTS/ --Title="AFR_vs_EUR_FitHiC_intersects"

#Compare across Ancestries for merged loops for both modalities
./Venn_Interactions.r --FileList=<path>/FULL_SAMPLES/loops/ALL_AFR_aggregated_merged_10kb_loops.bedpe,<path>/FULL_SAMPLES/loops/ALL_EUR_aggregated_merged_10kb_loops.bedpe --HeaderList=1,1 --Labels=AFRloops,EURloops --OutDir=<path>/FULL_SAMPLES/INTERSECTS/ --Title="AFR_vs_EUR_all_intersects_10kb"

#Compare across Methods for all samples' merged loops at once
./Venn_Interactions.r --FileList=<path>/FULL_SAMPLES/loops/ALL_aggregated_merged_10kb_FitHiC_loops.bedpe,<path>/FULL_SAMPLES/loops/ALL_aggregated_merged_10kb_HICCUPS_loops.bedpe --HeaderList=1,1 --Labels=FitHiC,HICCUPS --OutDir=<path>/FULL_SAMPLES/INTERSECTS/ --Title="FitHiC_vs_HICCUPS_ALL_10kb"

### Comparisons of ALL merged loops across ancestries and methods at 10kb against each of the E2G and HiChIP samples:
for FILE in Endothelium.EnhancerPredictionsThresholded.bedpe Fibroblast.EnhancerPredictionsThresholded.bedpe Myeloid.EnhancerPredictionsThresholded.bedpe VSMC_encode_e2g_predictions_threshold.171.bedpe; do
  ./Venn_Interactions.r --FileList=<path>/FULL_SAMPLES/loops/ALL_ALL_aggregated_merged_10kb_loops.bedpe,<path>/FULL_SAMPLES/loops/MultiomeHiC_revisions/E2G/${FILE} --HeaderList=1,1 --Labels=HiCLoops,E2G_${FILE} --OutDir=<path>/FULL_SAMPLES/INTERSECTS/ --Title="ALL_vs_E2G_${FILE}_intersects"
done

for FILE in FITHiChIP_EC_5kb.interactions_FitHiC_Q0.01_MergeNearContacts_filteredwDistance.bed FITHiChIP_SMC_5kb.interactions_FitHiC_Q0.01_MergeNearContacts_filteredwDistance.bed; do
  ./Venn_Interactions.r --FileList=<path>/FULL_SAMPLES/loops/ALL_ALL_aggregated_merged_10kb_loops.bedpe,<path>/FULL_SAMPLES/loops/MultiomeHiC_revisions/HiChIP_bedpe/${FILE} --HeaderList=1,1 --Labels=HiCLoops,HiCHIP_${FILE} --OutDir=<path>/FULL_SAMPLES/INTERSECTS/ --Title="ALL_vs_HiChIP_${FILE}_intersects"
done

#Repeat for ONLY the loops shared between HiCCUPS and FitHiC:
for FILE in Endothelium.EnhancerPredictionsThresholded.bedpe Fibroblast.EnhancerPredictionsThresholded.bedpe Myeloid.EnhancerPredictionsThresholded.bedpe VSMC_encode_e2g_predictions_threshold.171.bedpe; do
  ./Venn_Interactions.r --FileList=<path>/FULL_SAMPLES/INTERSECTS/FitHiC_vs_HICCUPS_ALL_10kb_Shared_Loops.bed,<path>/FULL_SAMPLES/loops/MultiomeHiC_revisions/E2G/${FILE} --HeaderList=1,1 --Labels=HiCLoops,E2G_${FILE} --OutDir=<path>/FULL_SAMPLES/INTERSECTS/ --Title="SHARED_ALL_vs_E2G_${FILE}_intersects"
done

for FILE in FITHiChIP_EC_5kb.interactions_FitHiC_Q0.01_MergeNearContacts_filteredwDistance.bed FITHiChIP_SMC_5kb.interactions_FitHiC_Q0.01_MergeNearContacts_filteredwDistance.bed; do
  ./Venn_Interactions.r --FileList=<path>/FULL_SAMPLES/INTERSECTS/FitHiC_vs_HICCUPS_ALL_10kb_Shared_Loops.bed,<path>/FULL_SAMPLES/loops/MultiomeHiC_revisions/HiChIP_bedpe/${FILE} --HeaderList=1,1 --Labels=HiCLoops,HiCHIP_${FILE} --OutDir=<path>/FULL_SAMPLES/INTERSECTS/ --Title="SHARED_ALL_vs_HiChIP_${FILE}_intersects"
done
```

## Calculate inferred max resolutions using HiCRes at the given sequencing depth
```{r}
#Get all sample IDs in a vector
samples <- c(SAMPLE_01, SAMPLE_02, SAMPLE_03, SAMPLE_04, SAMPLE_05, SAMPLE_06, SAMPLE_07, SAMPLE_08, SAMPLE_09, SAMPLE_10, SAMPLE_11, SAMPLE_12, SAMPLE_13, SAMPLE_14, SAMPLE_15, SAMPLE_16, SAMPLE_17, SAMPLE_18, SAMPLE_19, SAMPLE_20, SAMPLE_21, SAMPLE_22, "SAMPLE_05_old")

# Set up a function
reso_resolver <- function(SAMP){
  vars <- fread(paste0("<path>/FULL_SAMPLES/hicres_results/", SAMP, "_HiCRes_outputs/equation.txt"), skip=1)
  a <- vars$V3[1]
  b <- vars$V3[2]
  c <- vars$V3[3]
  d <- vars$V3[4]
  READS <- fread(list.files(paste0("<path>/FULL_SAMPLES/hicres_results/QC_outputs/", SAMP, "_QC_outputs/"), pattern="\\.txt$", full.names=TRUE)[1]) %>% dplyr::select(Valid_pairs)

  
  #Fromula provided by HiCRes paper (note formula as given in file is off, parentheses broken!)
  result <- (((1000)^(1/3) - d - c * (READS)^(1/3))/( a * (READS)^(1/3) + b)) ^ 3
  print(paste0(SAMP, " is at ", result, " on the basis of ", READS, " valid read pairs."))
}

#Iterate across all samples to get max inferred resolution
for (SAMP in samples) {
  reso_resolver(SAMP)  
}
```

## Merge significant loops identified by FitHiC:
```{r}
# Set working directory
setwd("<path>/FULL_SAMPLES/")

# List of samples and resolutions, as well as individual sample groupings
samples <- c(SAMPLE_01, SAMPLE_02, SAMPLE_03, SAMPLE_04, SAMPLE_05, SAMPLE_06, SAMPLE_07, SAMPLE_08, SAMPLE_09, SAMPLE_10, SAMPLE_11, SAMPLE_12, SAMPLE_13, SAMPLE_14, SAMPLE_15, SAMPLE_16, SAMPLE_17, SAMPLE_18, SAMPLE_19, SAMPLE_20, SAMPLE_21, SAMPLE_22)
coronary <- c(SAMPLE_01, SAMPLE_05, SAMPLE_09, SAMPLE_10, SAMPLE_11, SAMPLE_12, SAMPLE_13, SAMPLE_14, SAMPLE_15, SAMPLE_16, SAMPLE_17, SAMPLE_18, SAMPLE_19, SAMPLE_20, SAMPLE_21, SAMPLE_22)
cardiomyocyte <- c(SAMPLE_01, SAMPLE_05)
endothelial <- c(SAMPLE_02, SAMPLE_06)
fibroblast <- c(SAMPLE_03, SAMPLE_07)
macrophage <- c(SAMPLE_04, SAMPLE_08)
resolutions <- c(2000, 5000, 10000)

header <- c("chr1", "fragmentMid1", "chr2", "fragmentMid2", "contactCount", "p-value", "q-value", "bias1", "bias2", "ExpCC")

# Define the processing function
process_loops <- function(sample_group, group_name, resolutions) {
  for (res in resolutions) {
    all_loops <- list()
    
    # Loop over each sample in the group
    for (sample in sample_group) {
      
      # Construct file path for the filtered significant loops file
      file_path <- paste0("./", sample, "/fithic/", res, "/FitHiC.spline_pass1.res", res, ".significances.filtered.txt")
      
      # Check if the file exists
      if (file.exists(file_path)) {
        # Read in filtered loop file & provide appropriate column names
        test <- fread(file_path, header = FALSE, data.table = FALSE)
        colnames(test) <- header
        
        # Ensure correct data types
        test$chr1 <- as.character(test$chr1)
        test$fragmentMid1 <- as.numeric(test$fragmentMid1)
        test$chr2 <- as.character(test$chr2)
        test$fragmentMid2 <- as.numeric(test$fragmentMid2)
        
        # Diagnose problematic entries
        problematic_contact_count <- test$contactCount[!grepl("^[0-9]+(\\.[0-9]+)?$", test$contactCount)]
        if (length(problematic_contact_count) > 0) {
          message(paste("Non-numeric contactCount values found in sample", sample, "and resolution", res, ":", problematic_contact_count))
        }
        
        # Handle non-numeric values in contactCount by converting them to NA
        test$contactCount <- as.numeric(test$contactCount)
        test$`q-value` <- as.numeric(test$`q-value`)
        test$ExpCC <- as.numeric(test$ExpCC)
        
        # Create a unique ID for each loop
        test$ID <- paste(test$chr1, test$fragmentMid1, test$fragmentMid2, sep = "-")
        
        # Annotate with the sample ID
        test$Samples <- as.character(sample)
        
        # Append to the list of loops for this resolution
        all_loops[[length(all_loops) + 1]] <- test
      } else {
        message(paste("File not found for sample", sample, "and resolution", res))
      }
    }
    
    # Combine all data frames for this resolution into one
    combined_loops <- bind_rows(all_loops)
    
    # Ensure Samples column is character type before summarising
    combined_loops$Samples <- as.character(combined_loops$Samples)
    
    # Collapse to union list of all loops, merging loops that are the same across samples
    collapsed_loops <- combined_loops %>%
      group_by(ID) %>%
      summarise(
        chr1 = unique(chr1),
        fragmentMid1 = unique(fragmentMid1),
        fragmentMid2 = unique(fragmentMid2),
        contactCount = sum(contactCount, na.rm = TRUE),  # Sum of contact counts from different samples
        ExpCC = sum(ExpCC, na.rm = TRUE),  # Sum of ExpCC from different samples
        `q-value` = min(`q-value`, na.rm = TRUE),  # Keep the most significant q-value
        Samples = paste(Samples, collapse = ",")
      )
    
    # Convert to BEDPE format
    bedpe_loops <- collapsed_loops %>%
      mutate(
        start1 = fragmentMid1 - (res / 2),
        end1 = fragmentMid1 + (res / 2),
        start2 = fragmentMid2 - (res / 2),
        end2 = fragmentMid2 + (res / 2),
        name = ID,
        score = `q-value`,
        strand1 = ".",
        strand2 = "."
      ) %>%
      select(
        chr1, start1, end1, chr1, start2, end2, name, score, strand1, strand2,
        Samples, contactCount, ExpCC
      )
    
    bedpe_loops$total_samp <- str_count(bedpe_loops$Samples, ",")
    
    # Format the start and end columns to avoid scientific notation
    bedpe_loops$start1 <- format(bedpe_loops$start1, scientific = FALSE)
    bedpe_loops$end1 <- format(bedpe_loops$end1, scientific = FALSE)
    bedpe_loops$start2 <- format(bedpe_loops$start2, scientific = FALSE)
    bedpe_loops$end2 <- format(bedpe_loops$end2, scientific = FALSE)
    
    #Filter to loops found in >1 sample
    filtloops <- bedpe_loops[which(bedpe_loops$total_samp>=1),]
    fwrite(select(filtloops, -total_samp), paste0(group_name, "_merged_filt_loops_", res, ".bedpe"), sep="\t", quote=F, row.names=F, col.names=T)
    
    # Save the BEDPE result for this group and resolution to a file
    output_file <- paste0(group_name, "_loops_", res, ".bedpe")
    write.table(bedpe_loops, output_file, sep = "\t", row.names = FALSE, quote = FALSE, col.names = TRUE)
    
    message(paste("Finished processing", group_name, "at resolution", res))
  }
}

# Apply the function to different sample groupings
process_loops(samples, "ALL", resolutions)
process_loops(coronary, "coronary", resolutions)
process_loops(cardiomyocyte, "cardiomyocyte", resolutions)
process_loops(endothelial, "endothelial", resolutions)
process_loops(fibroblast, "fibroblast", resolutions)
process_loops(macrophage, "macrophage", resolutions)
```

## Look at all fithic significant loop calls, both checking on total #s and assessing overlaps:
```{r}
#Sample listing
samples <- c(SAMPLE_01, SAMPLE_02, SAMPLE_03, SAMPLE_04, SAMPLE_05, SAMPLE_06, SAMPLE_07, SAMPLE_08, SAMPLE_09, SAMPLE_10, SAMPLE_11, SAMPLE_12, SAMPLE_13, SAMPLE_14, SAMPLE_15, SAMPLE_16, SAMPLE_17, SAMPLE_18, SAMPLE_19, SAMPLE_20, SAMPLE_21, SAMPLE_22)

#Function to count individual loops
sigloop_counter <- function(sample, res) {
  myloops <- fread(paste0("<path>/FULL_SAMPLES/", sample, "/fithic/", res, "/FitHiC.spline_pass1.res", res, ".significances.filtered.txt"))
  print(paste0("Before filtering out self-loops, ", sample, " has ", nrow(myloops), " significant loops at ", res, " bp."))
  myloops <- myloops[V2 != V4, ]
  print(paste0("After filtering out self-loops, ", sample, " has ", nrow(myloops), " significant loops at ", res, " bp."))
}

#Count individual loops
lapply(samples, sigloop_counter, res=10000)
lapply(samples, sigloop_counter, res=5000)
lapply(samples, sigloop_counter, res=2000)

#Function to count merged loops, as well as visually assessing and then potentially providing a filtering based on a certain number of samples
mergeloop_counter <- function(grouping, res, n){
  myloops <- fread(paste0("<path>/FULL_SAMPLES/", grouping, "_loops_", res, ".bedpe"))
  print(paste0("Before filtering out self-loops, ", grouping, " has ", nrow(myloops), " significant loops at ", res, " bp."))
  myloops <- myloops[start1 != start2, ]
  print(paste0("After filtering out self-loops, ", grouping, " has ", nrow(myloops), " significant loops at ", res, " bp."))
  
  comma_counts <- str_count(myloops$Samples, ",")
  mycounts <- table(comma_counts)
  

  plotcounts <- data.frame(identified=as.numeric(names(mycounts))+1, loops=as.numeric(mycounts))
  p <- ggplot(plotcounts, aes(x=identified, y=loops)) + geom_bar(stat="identity") + xlab("# of Samples Inferring Loop") + ylab("Significant Loops") + ggtitle(paste0("Loop Counts X Samples Where Significant, ", grouping, " Loops, ", res/1000, "kb"))  +
  scale_x_continuous(breaks = seq(min(plotcounts$identified), max(plotcounts$identified), by = 1))
  print(p)
  print(plotcounts)
  plotcounts <- plotcounts[which(plotcounts$identified>=n),]
  print(sum(plotcounts$loops))
}

mergeloop_counter("ALL", 10000, 11)
mergeloop_counter("ALL", 5000, 11)
mergeloop_counter("ALL", 2000, 11)

mergeloop_counter("coronary", 10000, 7)
mergeloop_counter("coronary", 5000, 7)
mergeloop_counter("coronary", 2000, 7)

mergeloop_counter("cardiomyocyte", 10000, 2)
mergeloop_counter("cardiomyocyte", 5000, 2)
mergeloop_counter("cardiomyocyte", 2000, 2)

mergeloop_counter("endothelial", 10000, 2)
mergeloop_counter("endothelial", 5000, 2)
mergeloop_counter("endothelial", 2000, 2)

mergeloop_counter("fibroblast", 10000, 2)
mergeloop_counter("fibroblast", 5000, 2)
mergeloop_counter("fibroblast", 2000, 2)

mergeloop_counter("macrophage", 10000, 2)
mergeloop_counter("macrophage", 5000, 2)
mergeloop_counter("macrophage", 2000, 2)
```

## Write all significant loops into a reasonable BEDPE format for plotting:
```{r}
# Set working directory
setwd("<path>/FULL_SAMPLES/")

# List of samples and resolutions, as well as individual sample groupings
samples <- c(SAMPLE_01, SAMPLE_02, SAMPLE_03, SAMPLE_04, SAMPLE_05, SAMPLE_06, SAMPLE_07, SAMPLE_08, SAMPLE_09, SAMPLE_10, SAMPLE_11, SAMPLE_12, SAMPLE_13, SAMPLE_14, SAMPLE_15, SAMPLE_16, SAMPLE_17, SAMPLE_18, SAMPLE_19, SAMPLE_20, SAMPLE_21, SAMPLE_22)
resolutions <- c(2000, 5000, 10000)

header <- c("chr1", "fragmentMid1", "chr2", "fragmentMid2", "contactCount", "p-value", "q-value", "bias1", "bias2", "ExpCC")

BEDPE_formatter <- function(sampleID, resolution) {
        test <- fread(paste0("<path>/FULL_SAMPLES/", sampleID, "/fithic/", resolution, "/FitHiC.spline_pass1.res", resolution, ".significances.filtered.txt"), header = FALSE, data.table = FALSE)
        colnames(test) <- header
        
        # Ensure correct data types
        test$chr1 <- as.character(test$chr1)
        test$fragmentMid1 <- as.numeric(test$fragmentMid1)
        test$chr2 <- as.character(test$chr2)
        test$fragmentMid2 <- as.numeric(test$fragmentMid2)
        
        # Diagnose problematic entries
        problematic_contact_count <- test$contactCount[!grepl("^[0-9]+(\\.[0-9]+)?$", test$contactCount)]
        if (length(problematic_contact_count) > 0) {
          message(paste("Non-numeric contactCount values found in sample", sample, "and resolution", res, ":", problematic_contact_count))
        }
        
        # Handle non-numeric values in contactCount by converting them to NA
        test$contactCount <- as.numeric(test$contactCount)
        test$`q-value` <- as.numeric(test$`q-value`)
        test$ExpCC <- as.numeric(test$ExpCC)
        
        # Create a unique ID for each loop
        test$ID <- paste(test$chr1, test$fragmentMid1, test$fragmentMid2, sep = "-")
        
        # Annotate with the sample ID
        test$Samples <- as.character(sampleID)
        
        bedpe_loops <- test %>%
      mutate(
        start1 = fragmentMid1 - (resolution / 2),
        end1 = fragmentMid1 + (resolution / 2),
        start2 = fragmentMid2 - (resolution / 2),
        end2 = fragmentMid2 + (resolution / 2),
        name = ID,
        score = `q-value`,
        strand1 = ".",
        strand2 = "."
      ) %>%
      select(
        chr1, start1, end1, chr2, start2, end2, name, score, strand1, strand2,
        Samples, contactCount, ExpCC
      )
    
    # Format the start and end columns to avoid scientific notation
    bedpe_loops$start1 <- format(bedpe_loops$start1, scientific = FALSE)
    bedpe_loops$end1 <- format(bedpe_loops$end1, scientific = FALSE)
    bedpe_loops$start2 <- format(bedpe_loops$start2, scientific = FALSE)
    bedpe_loops$end2 <- format(bedpe_loops$end2, scientific = FALSE)
    
    # Save the BEDPE result for this group and resolution to a file
    output_file <- paste0("<path>/FULL_SAMPLES/", sampleID, "/fithic/", resolution, "/", sampleID, "_loops_", resolution, ".bedpe")
    write.table(bedpe_loops, output_file, sep = "\t", row.names = FALSE, quote = FALSE, col.names = TRUE)
}

#Iterate across all samples and resolutions to get bedpe files for each
for (SAMP in samples) {
  for(RES in resolutions){
    BEDPE_formatter(SAMP, RES)
  }
}
```

## Compute Jaccard Indices across samples at a given resolution, and heatmap cluster:
```{unix}
#!/bin/bash

# Define your samples and resolutions
SAMPLES=("SAMPLE_01" "SAMPLE_02" "SAMPLE_03" "SAMPLE_04" "SAMPLE_05" "SAMPLE_06" "SAMPLE_07" "SAMPLE_08")
RESOLUTIONS=("2000" "5000" "10000")

# Define the directory where your files are located
FILE_DIR="<path>/FULL_SAMPLES"

# Define the output file for Jaccard indices
OUTPUT_FILE="jaccard_indices.txt"

# Create the output file and write the header
echo -e "Samp1\tSamp2\tJaccard" > "$OUTPUT_FILE"

# Function to convert your data to bedPE format
convert_to_bedpe() {
    local infile=$1
    local outfile=$2
    local resolution=$3
    local half_window=$((resolution * 3 / 2))
    awk -v hw="$half_window" -F',' 'NR>1 {
        if ($1 != "chrM" && $3 != "chrM") {
            start1 = $2 - hw
            end1 = $2 + hw
            start2 = $4 - hw
            end2 = $4 + hw
            if (start1 < 0) start1 = 0
            if (start2 < 0) start2 = 0
            print $1 "\t" int(start1) "\t" int(end1) "\t" $3 "\t" int(start2) "\t" int(end2) "\tname\t.\t.\t."
        }
    }' "$infile" > "$outfile"
}

# Function to compute Jaccard index
compute_jaccard() {
    local file1=$1
    local file2=$2
    bedtools jaccard -a "$file1" -b "$file2" | awk 'NR==2 {print $3}'  # Get the Jaccard index value
}

# Iterate over samples and resolutions
for sample1 in "${SAMPLES[@]}"; do
    for resolution1 in "${RESOLUTIONS[@]}"; do
        file1="${FILE_DIR}/${sample1}/fithic/${resolution1}/${sample1}.sigloops.${resolution1}.txt"
        bedpe1="${FILE_DIR}/${sample1}/fithic/${resolution1}/${sample1}_${resolution1}.bedpe"
        convert_to_bedpe "$file1" "$bedpe1" "$resolution1"

        for sample2 in "${SAMPLES[@]}"; do
            for resolution2 in "${RESOLUTIONS[@]}"; do
                if [ "$sample1" != "$sample2" ] || [ "$resolution1" != "$resolution2" ]; then
                    file2="${FILE_DIR}/${sample2}/fithic/${resolution2}/${sample2}.sigloops.${resolution2}.txt"
                    bedpe2="${FILE_DIR}/${sample2}/fithic/${resolution2}/${sample2}_${resolution2}.bedpe"
                    convert_to_bedpe "$file2" "$bedpe2" "$resolution2"

                    jaccard_index=$(compute_jaccard "$bedpe1" "$bedpe2")
                    echo -e "${sample1}_${resolution1}\t${sample2}_${resolution2}\t$jaccard_index" >> "$OUTPUT_FILE"
                fi
            done
        done
    done
done
```

## Visualize Jaccard results
```{r}
setwd("<path>/FULL_SAMPLES/")
jaccard_data <- fread("jaccard_indices.txt")

#Add some metadata annotations
coronaries <- c(SAMPLE_09, SAMPLE_10, SAMPLE_11, SAMPLE_12, SAMPLE_13, SAMPLE_14, SAMPLE_15, SAMPLE_16, SAMPLE_17, SAMPLE_18, SAMPLE_19, SAMPLE_20, SAMPLE_21, SAMPLE_22, SAMPLE_05)
AA <- c(SAMPLE_09, SAMPLE_10, SAMPLE_11, SAMPLE_12, SAMPLE_13, SAMPLE_14, SAMPLE_15)
EUR <- c(SAMPLE_16, SAMPLE_17, SAMPLE_18, SAMPLE_19, SAMPLE_20, SAMPLE_21, SAMPLE_22, SAMPLE_05)
CM <- c(SAMPLE_01, SAMPLE_05)
ENDO <- c(SAMPLE_02, SAMPLE_06)
FIBRO <- c(SAMPLE_03, SAMPLE_07)
MACRO <- c(SAMPLE_04, SAMPLE_08)

jaccard_data$Samp1 <- ifelse(jaccard_data$Samp1 %in% coronaries, paste0(jaccard_data$Samp1, "_", "coronary"), jaccard_data$Samp1)
jaccard_data$Samp2 <- ifelse(jaccard_data$Samp2 %in% coronaries, paste0(jaccard_data$Samp2, "_", "coronary"), jaccard_data$Samp2)

jaccard_data$Samp1 <- ifelse(gsub("_.*", "", jaccard_data$Samp1) %in% AA, paste0(jaccard_data$Samp1, "_", "AA"), jaccard_data$Samp1)
jaccard_data$Samp2 <- ifelse(gsub("_.*", "", jaccard_data$Samp2) %in% AA, paste0(jaccard_data$Samp2, "_", "AA"), jaccard_data$Samp2)

jaccard_data$Samp1 <- ifelse(gsub("_.*", "", jaccard_data$Samp1) %in% EUR, paste0(jaccard_data$Samp1, "_", "EUR"), jaccard_data$Samp1)
jaccard_data$Samp2 <- ifelse(gsub("_.*", "", jaccard_data$Samp2) %in% EUR, paste0(jaccard_data$Samp2, "_", "EUR"), jaccard_data$Samp2)

jaccard_data$Samp1 <- ifelse(jaccard_data$Samp1 %in% CM, paste0(jaccard_data$Samp1, "_", "CM"), jaccard_data$Samp1)
jaccard_data$Samp2 <- ifelse(jaccard_data$Samp2 %in% CM, paste0(jaccard_data$Samp2, "_", "CM"), jaccard_data$Samp2)

jaccard_data$Samp1 <- ifelse(jaccard_data$Samp1 %in% ENDO, paste0(jaccard_data$Samp1, "_", "ENDO"), jaccard_data$Samp1)
jaccard_data$Samp2 <- ifelse(jaccard_data$Samp2 %in% ENDO, paste0(jaccard_data$Samp2, "_", "ENDO"), jaccard_data$Samp2)

jaccard_data$Samp1 <- ifelse(jaccard_data$Samp1 %in% FIBRO, paste0(jaccard_data$Samp1, "_", "FIBRO"), jaccard_data$Samp1)
jaccard_data$Samp2 <- ifelse(jaccard_data$Samp2 %in% FIBRO, paste0(jaccard_data$Samp2, "_", "FIBRO"), jaccard_data$Samp2)

jaccard_data$Samp1 <- ifelse(jaccard_data$Samp1 %in% MACRO, paste0(jaccard_data$Samp1, "_", "MACRO"), jaccard_data$Samp1)
jaccard_data$Samp2 <- ifelse(jaccard_data$Samp2 %in% MACRO, paste0(jaccard_data$Samp2, "_", "MACRO"), jaccard_data$Samp2)


#2kb
# Convert to a wide matrix format where rows and columns represent samples
jaccard_matrix <- reshape2::acast(jaccard_data[which(jaccard_data$Resolution==2000)], Samp1 ~ Samp2, value.var = "Jaccard")
# Ensure that the matrix is symmetric (optional, depending on your data)
jaccard_matrix[lower.tri(jaccard_matrix)] <- t(jaccard_matrix)[lower.tri(jaccard_matrix)]
jaccard_matrix[is.na(jaccard_matrix)] <- 1  # Replace NA values with 1 as they are for self-identity; this didn't work correctly on my first set of data as we also had some NAs across samples where a chromosome existed in one but not the other (specifically chrY, so I removed this from all analyses)

pheatmap(
  jaccard_matrix,
  cluster_rows = TRUE,   # Enable hierarchical clustering for rows
  cluster_cols = TRUE,   # Enable hierarchical clustering for columns
  scale = "none",        # No scaling by default (use "row" or "column" if desired)
  color = colorRampPalette(c("blue", "white", "red"))(50),  # Custom color scale
  main = "Heatmap with Hierarchical Clustering"
)

#5kb
# Convert to a wide matrix format where rows and columns represent samples
jaccard_matrix <- reshape2::acast(jaccard_data[which(jaccard_data$Resolution==5000)], Samp1 ~ Samp2, value.var = "Jaccard")
# Ensure that the matrix is symmetric (optional, depending on your data)
jaccard_matrix[lower.tri(jaccard_matrix)] <- t(jaccard_matrix)[lower.tri(jaccard_matrix)]
jaccard_matrix[is.na(jaccard_matrix)] <- 1  # Replace NA values with 1 as they are for self-identity; this didn't work correctly on my first set of data as we also had some NAs across samples where a chromosome existed in one but not the other (specifically chrY, so I removed this from all analyses)

pheatmap(
  jaccard_matrix,
  cluster_rows = TRUE,   # Enable hierarchical clustering for rows
  cluster_cols = TRUE,   # Enable hierarchical clustering for columns
  scale = "none",        # No scaling by default (use "row" or "column" if desired)
  color = colorRampPalette(c("blue", "white", "red"))(50),  # Custom color scale
  main = "Heatmap with Hierarchical Clustering"
)

#10kb
# Convert to a wide matrix format where rows and columns represent samples
jaccard_matrix <- reshape2::acast(jaccard_data[which(jaccard_data$Resolution==10000)], Samp1 ~ Samp2, value.var = "Jaccard")
# Ensure that the matrix is symmetric (optional, depending on your data)
jaccard_matrix[lower.tri(jaccard_matrix)] <- t(jaccard_matrix)[lower.tri(jaccard_matrix)]
jaccard_matrix[is.na(jaccard_matrix)] <- 1  # Replace NA values with 1 as they are for self-identity; this didn't work correctly on my first set of data as we also had some NAs across samples where a chromosome existed in one but not the other (specifically chrY, so I removed this from all analyses)

pheatmap(
  jaccard_matrix,
  cluster_rows = TRUE,   # Enable hierarchical clustering for rows
  cluster_cols = TRUE,   # Enable hierarchical clustering for columns
  scale = "none",        # No scaling by default (use "row" or "column" if desired)
  color = colorRampPalette(c("blue", "white", "red"))(50),  # Custom color scale
  main = "Heatmap with Hierarchical Clustering"
)
```

## Fix human GFF file to GTF file for Hi-C visualization
```{r}
setwd("<path>/FULL_SAMPLES/")

# Read the GTF file into a data frame
gtf <- read.table("hg38.refGene.gtf", sep="\t", header=FALSE, stringsAsFactors=FALSE, quote="")
gtf <- gtf[!grepl("_alt|_random|_fix|chrUn|chrM", gtf$V1, perl=TRUE),]

# Filter to include only transcripts!!
transcripts <- gtf %>% filter(V3 == "transcript")

# Extract gene_id from the attribute column (V9)
transcripts <- transcripts %>% mutate(gene_id = paste(V1, sub('.*gene_id "([^"]+)".*', '\\1', V9), sep=":"), transcript_id = sub('.*transcript_id "([^"]+)".*', '\\1', V9))

# Collapse transcripts into single gene entries
collapsed <- transcripts %>%
  group_by(transcript_id) %>%
  dplyr::summarize(
    chr=first(V1),
    start = min(V4),
    end = max(V5),
    strand = first(V7),
    gene_id = first(gene_id)
  )

# Create the collapsed GTF format
collapsed_gtf <- collapsed %>%
  dplyr::mutate(
    source = "collapsed",
    feature = "gene",
    score = ".",
    frame = ".",
    attribute = paste0('gene_id "', gsub("chr.*:", "", gene_id), '"; gene_source "ittai";')
  ) %>% select(chr, source, feature, start, end, score, strand, frame, attribute)

collapsed_gtf <- as.data.frame(collapsed_gtf) %>% arrange(chr, start) %>% distinct()

# Write to a new GTF file
write.table(collapsed_gtf, "hg38.refGene.collapsed.gtf", sep="\t", quote=FALSE, row.names=FALSE, col.names=FALSE)

#The above DID work but still leaves more copies per gene than I would like, try an alternative and hopefully simpler approach:
library(TxDb.Hsapiens.UCSC.hg38.knownGene)
library(org.Hs.eg.db)
library(AnnotationDbi)
library(dplyr)

txdb <- TxDb.Hsapiens.UCSC.hg38.knownGene

genes_df <- genes(txdb) %>%
  as.data.frame()

gene_symbols <- select(org.Hs.eg.db,
                       keys = genes_df$gene_id,
                       columns = c("SYMBOL"),
                       keytype = "ENTREZID")

# Merge the gene symbols with the genes dataframe
genes_df <- merge(genes_df, gene_symbols, by.x = "gene_id", by.y = "ENTREZID")

gtf_df <- genes_df %>%
  mutate(seqname = seqnames,
         source = "hg38",
         feature = "gene",
         start = start,
         end = end,
         score = ".",
         strand = strand,
         frame = ".",
         gene_id = SYMBOL) %>%
  dplyr::select(seqname, source, feature, start, end, score, strand, frame, gene_id) %>% arrange(seqname, start)

gtf_df <- gtf_df[-which(is.na(gtf_df$gene_id)),]
gtf_df$gene_id <- paste0(' gene_id "', gtf_df$gene_id, '"; gene_source "ittai";')

# Rename columns to match GTF format
colnames(gtf_df) <- c("seqname", "source", "feature", "start", "end", "score", "strand", "frame", "gene_id")
write.table(gtf_df, "hg38.custom.geneBodyOnly.gtf", sep="\t", quote=FALSE, row.names=FALSE, col.names=FALSE)
```

## Differential TADs using TADCompare
### First, find consensus TAD boundaries across ancestries:
```{r}
#Function to read-in data from HiC-Pro:
TAD_dat_reader <- function(path, res){
  bed <- read.table(paste0(path, ".", res, "_abs.bed"))
  bed[,1] <- paste0("chr", bed[,1])
  mat <- read.table(paste0(path, ".", res, ".matrix"))
  dat <- HiCcompare::hicpro2bedpe(mat, bed)
  dat <- dat$cis
  dat <- dat[-grep("Un|alt|random|X|Y|M", names(dat), perl=T)]
  
  sparse_mat <- lapply(1:length(dat), function(z){
    x <- dat[[z]]
    chr <- x[,1][1]
    x <- x[c(2,5,7)]
    return(list(x, chr))
  })
  return(sparse_mat)
}

SSAMPLE_16 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_16/raw/SAMPLE_16", 50000)
SSAMPLE_17 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_17/raw/SAMPLE_17", 50000)
SSAMPLE_18 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_18/raw/SAMPLE_18", 50000)
SSAMPLE_19 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_19/raw/SAMPLE_19", 50000)
SSAMPLE_20 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_20/raw/SAMPLE_20", 50000)
SSAMPLE_21 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_21/raw/SAMPLE_21", 50000)
SSAMPLE_22 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_22/raw/SAMPLE_22", 50000)
SSAMPLE_05 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_05/raw/SAMPLE_05", 50000)

SSAMPLE_09 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_09/raw/SAMPLE_09", 50000)
SSAMPLE_10 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_10/raw/SAMPLE_10", 50000)
SSAMPLE_11 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_11/raw/SAMPLE_11", 50000)
SSAMPLE_12 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_12/raw/SAMPLE_12", 50000)
SSAMPLE_13 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_13/raw/SAMPLE_13", 50000)
SSAMPLE_14 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_14/raw/SAMPLE_14", 50000)
SSAMPLE_15 <- TAD_dat_reader("<path>/FULL_SAMPLES/SAMPLE_15/raw/SAMPLE_15", 50000)

#This works, need to wrap it for length of the samples (22):
EUR_TAD <- ConsensusTADs(c(SSAMPLE_16[[21]][1], SSAMPLE_17[[21]][1], SSAMPLE_18[[21]][1], SSAMPLE_19[[21]][1], SSAMPLE_20[[21]][1], SSAMPLE_21[[21]][1], SSAMPLE_22[[21]][1], SSAMPLE_05[[21]][1]), resolution=50000)

results <- lapply(1:22, function(i) {
  ConsensusTADs(c(SSAMPLE_16[[i]][1], SSAMPLE_17[[i]][1], SSAMPLE_18[[i]][1], SSAMPLE_19[[i]][1], SSAMPLE_20[[i]][1], SSAMPLE_21[[i]][1], SSAMPLE_22[[i]][1], SSAMPLE_05[[i]][1]), resolution=50000)
})

# Extract the $Consensus element from each result and combine them into a list
consensus_list <- lapply(results, function(x) x$Consensus)

# Combine the list into a data frame
df <- do.call(rbind, consensus_list)

# Add a new column for the chromosome
df$chromosome <- rep(paste0("chr", 1:22), sapply(consensus_list, nrow))
fwrite(df, "consensus_EUR_TADs.txt", sep="\t", quote=F, row.names=F, col.names=T)

results_AFR <- lapply(1:22, function(i) {
  ConsensusTADs(c(SSAMPLE_09[[i]][1], SSAMPLE_10[[i]][1], SSAMPLE_11[[i]][1], SSAMPLE_12[[i]][1], SSAMPLE_13[[i]][1], SSAMPLE_14[[i]][1], SSAMPLE_15[[i]][1]), resolution=50000)
})

# Extract the $Consensus element from each result and combine them into a list
consensus_list_AFR <- lapply(results_AFR, function(x) x$Consensus)

# Combine the list into a data frame
df_AFR <- do.call(rbind, consensus_list_AFR)

# Add a new column for the chromosome
df_AFR$chromosome <- rep(paste0("chr", 1:22), sapply(consensus_list_AFR, nrow))

# Write the data frame to a file
fwrite(df_AFR, "consensus_AFR_TADs.txt", sep="\t", quote=F, row.names=F, col.names=T)
```


#Differential TAD testing
```{r}
# Function to read and format data
read_format_data <- function(path, res) {
  bed <- read.table(paste0(path, ".", res, "_abs.bed"))
  bed[,1] <- paste0("chr", bed[,1])
  mat <- read.table(paste0(path, ".", res, ".matrix"))
  dat <- HiCcompare::hicpro2bedpe(mat, bed)
  dat <- dat$cis
  dat <- dat[-grep("Un|alt|random|X|Y|M", names(dat), perl=T)]
  return(dat)
}

# Function to make pairwise comparisons
compare_pairwise <- function(dat1, dat2) {
  sparse_tads = lapply(1:length(dat1), function(z) {
    x <- dat1[[z]]
    y <- dat2[[z]]
    
    # Pull out chromosome
    chr <- x[, 1][1]
    # Subset to make three column matrix
    x <- x[, c(2, 5, 7)]
    y <- y[, c(2, 5, 7)]
    
    # Run SpectralTAD
    comp <- TADCompare(x, y, resolution = 50000)
    return(list("comp" = comp, "chr" = chr))
  })
  
  diff_res <- lapply(sparse_tads, function(x) x[["comp"]])
  chr <- lapply(sparse_tads, function(x) x[["chr"]])
  names(diff_res) <- chr
  return(diff_res)
}

# Example usage
#dat1 <- read_format_data(path1, res)
#dat2 <- read_format_data(path2, res)
#result <- compare_pairwise(dat1, dat2)

# List of paths to the AFR and EUR samples
AFR_paths <- c("<path>/FULL_SAMPLES/SAMPLE_09/raw/SAMPLE_09", 
               "<path>/FULL_SAMPLES/SAMPLE_10/raw/SAMPLE_10", 
               "<path>/FULL_SAMPLES/SAMPLE_11/raw/SAMPLE_11", 
               "<path>/FULL_SAMPLES/SAMPLE_12/raw/SAMPLE_12", 
               "<path>/FULL_SAMPLES/SAMPLE_13/raw/SAMPLE_13", 
               "<path>/FULL_SAMPLES/SAMPLE_14/raw/SAMPLE_14", 
               "<path>/FULL_SAMPLES/SAMPLE_15/raw/SAMPLE_15")
EUR_paths <- c("<path>/FULL_SAMPLES/SAMPLE_16/raw/SAMPLE_16", 
               "<path>/FULL_SAMPLES/SAMPLE_17/raw/SAMPLE_17", 
               "<path>/FULL_SAMPLES/SAMPLE_18/raw/SAMPLE_18", 
               "<path>/FULL_SAMPLES/SAMPLE_19/raw/SAMPLE_19", 
               "<path>/FULL_SAMPLES/SAMPLE_20/raw/SAMPLE_20", 
               "<path>/FULL_SAMPLES/SAMPLE_21/raw/SAMPLE_21", 
               "<path>/FULL_SAMPLES/SAMPLE_22/raw/SAMPLE_22", 
               "<path>/FULL_SAMPLES/SAMPLE_05/raw/SAMPLE_05")

# Extract the sample number IDs from the paths
AFR_ids <- basename(AFR_paths)
EUR_ids <- basename(EUR_paths)

# Load the data for each sample
AFR_samples <- lapply(AFR_paths, read_format_data, res = 50000)
EUR_samples <- lapply(EUR_paths, read_format_data, res = 50000)

# Initialize an empty list to store the results
results <- list()

# Loop over all AFR and EUR samples
for (i in seq_along(AFR_samples)) {
  for (j in seq_along(EUR_samples)) {
    # Run compare_pairwise on the current pair of samples
    result <- compare_pairwise(AFR_samples[[i]], EUR_samples[[j]])
    # Add the result to the list, with a name indicating the pair of samples
    results[[paste0("AFR", AFR_ids[i], "_EUR", EUR_ids[j])]] <- result
  }
}


#Subset down to what we care about, $TAD_Frame, the first entry in each list entry that actually lists out the differential TAD boundaries (second entry lists ALL tested bins, first retains only those that are a boundary in at least one matrix). We really only care about these potential regions of difference, so just extract the first data frame in the list. The second will be similar but at all tested bins (most of which won't be a boundary in either sample!), and the last item is a ggplot2 stacked bar chart of a table of the "Type" column from the first data frame, so can be easily reconstructed (and indeed, I will want to, since it is done per-sample and per-chromosome now, but I will want at least merged across all chromosomes, if not merged across sample groups as well).
# Initialize an empty list to store the data frames
df_list <- list()

# Iterate over the results list
for (comparison_name in names(results)) {
  comparison_results <- results[[comparison_name]]
  
  # Iterate over the chromosomes in each comparison result
  for (chromosome_name in names(comparison_results)) {
    chromosome_result <- comparison_results[[chromosome_name]]
    
    # Extract the TAD_Frame data frame and add the chromosome and comparison columns
    df <- chromosome_result$TAD_Frame
    df$Chromosome <- chromosome_name
    df$Comparison <- comparison_name
    
    # Add the data frame to the list
    df_list[[paste0(comparison_name, "_", chromosome_name)]] <- df
  }
}

# Combine all the data frames into a single data frame
final_df <- do.call(rbind, df_list)

fwrite(final_df, "Final_differential_TAD_bounds.txt", sep="\t", quote=F, row.names=F, col.names=T)






#Take a look at intra-ancestry comparisons:
# Initialize an empty list to store the results_intra
results_intra <- list()
# Loop over all AFR samples for intra-group comparisons
for (i in seq_along(AFR_samples)) {
  for (j in i:seq_along(AFR_samples)) {
    if (i != j) {
      # Run compare_pairwise on the current pair of samples
      result <- compare_pairwise(AFR_samples[[i]], AFR_samples[[j]])
      # Add the result to the list, with a name indicating the pair of samples
      results_intra[[paste0("AFR", AFR_ids[i], "_AFR", AFR_ids[j])]] <- result
    }
  }
}

# Loop over all EUR samples for intra-group comparisons
for (i in seq_along(EUR_samples)) {
  for (j in i:seq_along(EUR_samples)) {
    if (i != j) {
      # Run compare_pairwise on the current pair of samples
      result <- compare_pairwise(EUR_samples[[i]], EUR_samples[[j]])
      # Add the result to the list, with a name indicating the pair of samples
      results_intra[[paste0("EUR", EUR_ids[i], "_EUR", EUR_ids[j])]] <- result
    }
  }
}

df_list <- list()

# Iterate over the results_intra list
for (comparison_name in names(results_intra)) {
  comparison_results_intra <- results_intra[[comparison_name]]
  
  # Iterate over the chromosomes in each comparison result
  for (chromosome_name in names(comparison_results_intra)) {
    chromosome_result <- comparison_results_intra[[chromosome_name]]
    
    # Extract the TAD_Frame data frame and add the chromosome and comparison columns
    df <- chromosome_result$TAD_Frame
    df$Chromosome <- chromosome_name
    df$Comparison <- comparison_name
    
    # Add the data frame to the list
    df_list[[paste0(comparison_name, "_", chromosome_name)]] <- df
  }
}

# Combine all the data frames into a single data frame
final_df <- do.call(rbind, df_list)

fwrite(final_df, "Final_intraancestry_differential_TAD_bounds.txt", sep="\t", quote=F, row.names=F, col.names=T)
```

#Now, examine results. Ideally, want to visualize some pairwise comparisons at differential TADs that overlap the consensus TADs called previously. Load in the consensus TADs from prior run first:
```{r}
setwd("<path>")
AFR.cons <- fread("consensus_AFR_TADs.txt")
EUR.cons <- fread("consensus_EUR_TADs.txt")
diff.TADs <- fread("Final_differential_TAD_bounds.txt")
intra.diff.TADs <- fread("Final_intraancestry_differential_TAD_bounds.txt")

AFR.cons$boundID <- paste(AFR.cons$chromosome, AFR.cons$Coordinate, sep=":")
EUR.cons$boundID <- paste(EUR.cons$chromosome, EUR.cons$Coordinate, sep=":")
diff.TADs$boundID <- paste(diff.TADs$Chromosome, diff.TADs$Boundary, sep=":")
intra.diff.TADs$boundID <- paste(intra.diff.TADs$Chromosome, intra.diff.TADs$Boundary, sep=":")

sum(AFR.cons$boundID %in% EUR.cons$boundID)
sum(EUR.cons$boundID %in% AFR.cons$boundID)

sum(EUR.cons$boundID %in% diff.TADs$boundID)
sum(AFR.cons$boundID %in% diff.TADs$boundID)

sum(diff.TADs$boundID %in% intra.diff.TADs$boundID)
sum(intra.diff.TADs$boundID %in% diff.TADs$boundID)

sum(diff.TADs$boundID %in% AFR.cons$boundID | diff.TADs$boundID %in% EUR.cons$boundID)
signif.TADs <- diff.TADs[which(diff.TADs$boundID %in% AFR.cons$boundID | diff.TADs$boundID %in% EUR.cons$boundID),]

#Take a look at TAD boundaries across comparisons
ggplot(filter(diff.TADs, Type!=""), aes(x="Type", fill=Type)) + geom_bar() + facet_wrap(~Comparison) + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(y="Boundary Count", fill="Type") + ggtitle("TAD Boundary Comparisons")

#Generate a single barchart for Brandon's CMD presentation October 2024:
diff.TADs$Type <- factor(diff.TADs$Type, levels=c("Split", "Complex", "Merge", "Strength Change", "Shifted", "Non-Differential"))
ggplot(filter(diff.TADs, Type!="", Comparison=="AFRSAMPLE_09_EURSAMPLE_16"), aes(x="Type", fill=Type)) + geom_bar() + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(y="Boundary Count", fill="Type") + ggtitle("TAD Boundary Comparisons") + scale_fill_manual(values=c("Split"="#00BFC4", "Complex"="#F8766D", "Merge"="#B79F00", "Strength Change"="#F564E3", "Shifted"="#619CFF", "Non-Differential"="#00BA38"))


ggplot(filter(diff.TADs, Type!="Non-Differential"&Type!=""), aes(x="Type", fill=Type)) + geom_bar() + facet_wrap(~Comparison) + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(y="Boundary Count", fill="Type") + ggtitle("TAD Boundary Comparisons")

ggplot(filter(intra.diff.TADs, Type!=""), aes(x="Type", fill=Type)) + geom_bar() + facet_wrap(~Comparison) + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(y="Boundary Count", fill="Type") + ggtitle("TAD Boundary Comparisons")

ggplot(filter(intra.diff.TADs, Type!="Non-Differential"&Type!=""), aes(x="Type", fill=Type)) + geom_bar() + facet_wrap(~Comparison) + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(y="Boundary Count", fill="Type") + ggtitle("TAD Boundary Comparisons")

#Get numbers for these comparisons
table(intra.diff.TADs[grep("EUR", intra.diff.TADs$Comparison),]$Type)
table(intra.diff.TADs[grep("AFR", intra.diff.TADs$Comparison),]$Type)
table(diff.TADs$Type)

table(diff.TADs[-grep("SAMPLE_09|SAMPLE_10|SAMPLE_11|SAMPLE_16|SAMPLE_17|SAMPLE_18|SAMPLE_05", diff.TADs$Comparison),]$Type) #AFRm vs EURm; remove females for this
table(diff.TADs[-grep("SAMPLE_12|SAMPLE_13|SAMPLE_14|SAMPLE_15|SAMPLE_19|SAMPLE_20|SAMPLE_21|SAMPLE_22", diff.TADs$Comparison),]$Type) #AFRf vs EURf; remove males for this

#Look across sexes within ancestries.
allcomps <- unique(intra.diff.TADs$Comparison)
AFRmAFRf <- paste(allcomps[c(4:6,8:10,13:15,19:21)], collapse="|")
AFRfs <- paste(allcomps[c(1:3)], collapse="|")
AFRms <- paste(allcomps[c(7,11,12,16,17)], collapse="|")
table(intra.diff.TADs[grep(AFRmAFRf, intra.diff.TADs$Comparison),]$Type) #AFRm vs AFRf, have to remove any AFRm-AFRm and AFRf-AFRf comparisons for this
table(intra.diff.TADs[grep(AFRfs, intra.diff.TADs$Comparison),]$Type) #AFRfs
table(intra.diff.TADs[grep(AFRms, intra.diff.TADs$Comparison),]$Type) #AFRms

EURmEURf <- paste(allcomps[c(25:27,29:31,34:36,40:46)], collapse="|")
EURfs <- paste(allcomps[c(22:24,47:49)], collapse="|")
EURms <- paste(allcomps[c(28,32:33,37:39)], collapse="|")

table(intra.diff.TADs[grep(EURmEURf, intra.diff.TADs$Comparison),]$Type) #EURm vs EURf, have to remove any EURm-EURm and EURf-EURf comparisons for this
table(intra.diff.TADs[grep(EURfs, intra.diff.TADs$Comparison),]$Type) #EURfs
table(intra.diff.TADs[grep(EURms, intra.diff.TADs$Comparison),]$Type) #EURms

#Gap score is the differential boundary score, subset to boundaries with strong change and see which pop up most:
sort(table(diff.TADs[which(abs(diff.TADs$Gap_Score)>=15),]$boundID), decreasing=T)

#Several potentially compelling candidates emerge: chr6:5.9e+07, chr2:92750000, chr2:93100000 are the strongest, off the cuff.
diff.TADs[which(diff.TADs$boundID=="chr6:5.9e+07"),]
diff.TADs[which(diff.TADs$boundID=="chr2:92750000"),]
diff.TADs[which(diff.TADs$boundID=="chr2:93100000"),]

#Next level down of potential candidates, with 9 hits from above filtering, includes: chr2:9.3e+07, chr2:92900000, chr3:92750000, chr4:5.1e+07, chr5:49450000. These all have 9 hits, then goes down to 8.



#Look only at the consensus TAD boundaries:
#Take a look at TAD boundaries across comparisons
ggplot(filter(diff.TADs, Type!="", boundID %in% EUR.cons$boundID|boundID %in% AFR.cons$boundID), aes(x="Type", fill=Type)) + geom_bar() + facet_wrap(~Comparison) + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(y="Boundary Count", fill="Type") + ggtitle("TAD Boundary Comparisons")

ggplot(filter(diff.TADs, Type!="Non-Differential"&Type!="", boundID %in% EUR.cons$boundID|boundID %in% AFR.cons$boundID), aes(x="Type", fill=Type)) + geom_bar() + facet_wrap(~Comparison) + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(y="Boundary Count", fill="Type") + ggtitle("TAD Boundary Comparisons")

ggplot(filter(diff.TADs, Type!="Non-Differential"&Type!="", boundID %in% EUR.cons$boundID&boundID %nin% AFR.cons$boundID), aes(x="Type", fill=Type)) + geom_bar() + facet_wrap(~Comparison) + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(y="Boundary Count", fill="Type") + ggtitle("TAD Boundary Comparisons")

ggplot(filter(diff.TADs, Type!="Non-Differential"&Type!="", boundID %nin% EUR.cons$boundID&boundID %in% AFR.cons$boundID), aes(x="Type", fill=Type)) + geom_bar() + facet_wrap(~Comparison) + theme(axis.text.x=element_text(angle=90, hjust=1)) + labs(y="Boundary Count", fill="Type") + ggtitle("TAD Boundary Comparisons")
```

### Tissue based differential TADs
```{r}
#Below code was from the ancestry example, now, we want to compare across tissues as follows:
#Endo: SAMPLE_02 and SAMPLE_06
#Fibro: SAMPLE_03 and SAMPLE_07
#Macro: SAMPLE_04 and SAMPLE_08
#CA: SAMPLE_01 and SAMPLE_05


# Required library
library(HiCcompare)
library(dplyr)
library(data.table)
library(HiCcompare)
library(multiHiCcompare)
library(SpectralTAD)
library(TADCompare)
library(BiocParallel)
library(ggplot2)
library(Hmisc)
numCores <- 96
register(MulticoreParam(workers=numCores), default=TRUE)
data("hg38_cyto")

# Function to read and format data
read_format_data <- function(id) {
  path <- paste0("<path>/FULL_SAMPLES/", id, "/raw/", id)
  res <- "50000"
  bed <- read.table(paste0(path, ".", res, "_abs.bed"))
  bed[,1] <- paste0("chr", bed[,1])
  mat <- read.table(paste0(path, ".", res, ".matrix"))
  dat <- HiCcompare::hicpro2bedpe(mat, bed)
  dat <- dat$cis
  dat <- dat[-grep("Un|alt|random|X|Y|M", names(dat), perl=TRUE)]
  return(dat)
}

# Function to make pairwise comparisons
compare_pairwise <- function(dat1, dat2) {
  sparse_tads = lapply(1:length(dat1), function(z) {
    x <- dat1[[z]]
    y <- dat2[[z]]
    chr <- x[, 1][1]  # Pull out chromosome
    x <- x[, c(2, 5, 7)]  # Subset to make three column matrix
    y <- y[, c(2, 5, 7)]
    comp <- TADCompare(x, y, resolution = 50000)
    return(list("comp" = comp, "chr" = chr))
  })
  diff_res <- lapply(sparse_tads, function(x) x[["comp"]])
  chr <- lapply(sparse_tads, function(x) x[["chr"]])
  names(diff_res) <- chr
  return(diff_res)
}

# Sample IDs for each tissue type
sample_ids <- list(
  Endo = c("SAMPLE_02", "SAMPLE_06"),
  Fibro = c("SAMPLE_03", "SAMPLE_07"),
  Macro = c("SAMPLE_04", "SAMPLE_08"),
  CA = c("SAMPLE_01", "SAMPLE_05")
)

# Load the data for each sample
samples_data <- lapply(sample_ids, function(ids) {
  sapply(ids, read_format_data, simplify = FALSE)
})

# Intra-tissue comparisons
intra_tissue_results <- lapply(samples_data, function(tissue_samples) {
  results <- list()
  for (i in 1:(length(tissue_samples) - 1)) {
    for (j in (i + 1):length(tissue_samples)) {
      result <- compare_pairwise(tissue_samples[[i]], tissue_samples[[j]])
      results[[paste0(names(tissue_samples)[i], "_vs_", names(tissue_samples)[j])]] <- result
    }
  }
  return(results)
})

# Inter-tissue comparisons
inter_tissue_results <- list()
tissue_names <- names(samples_data)
for (i in 1:(length(samples_data) - 1)) {
  for (j in (i + 1):length(samples_data)) {
    for (sample_i in samples_data[[i]]) {
      for (sample_j in samples_data[[j]]) {
        comparison_name <- paste0(names(sample_i), "_vs_", names(sample_j))
        result <- compare_pairwise(sample_i, sample_j)
        inter_tissue_results[[comparison_name]] <- result
      }
    }
  }
}

# Inter-tissue comparisons
inter_tissue_results <- list()
tissue_names <- names(samples_data)

# Adjusting loop to correctly handle sample data and naming
for (i in 1:(length(samples_data) - 1)) {
  for (j in (i + 1):length(samples_data)) {
    tissue_i_names <- names(samples_data[[tissue_names[i]]])
    tissue_j_names <- names(samples_data[[tissue_names[j]]])
    
    for (k in 1:length(tissue_i_names)) {
      for (l in 1:length(tissue_j_names)) {
        sample_i <- samples_data[[tissue_names[i]]][[tissue_i_names[k]]]
        sample_j <- samples_data[[tissue_names[j]]][[tissue_j_names[l]]]
        
        comparison_name <- paste0(tissue_i_names[k], "_vs_", tissue_j_names[l])
        
        result <- compare_pairwise(sample_i, sample_j)
        inter_tissue_results[[comparison_name]] <- result
      }
    }
  }
}


# At this point, `intra_tissue_results` contains all the comparisons within each tissue type,
# and `inter_tissue_results` contains comparisons across different tissue types.
# You can further process these results as needed, for example, by extracting TAD differences
# or visualizing the comparison outcomes.
# Iterate over the results list
df_list <- list()
for (comparison_name in names(intra_tissue_results)) {
  comparison_results <- intra_tissue_results[[comparison_name]]
  
  # Add the data frame to the list
  df_list[[paste0(comparison_name, "_", chromosome_name)]] <- comparison_results
}


# Combine all the data frames into a single data frame
intra_df <- do.call(rbind, df_list)
fwrite(intra_df, "Intra-tissue_differential_TAD_bounds.txt", sep="\t", quote=F, row.names=F, col.names=T)


df_list <- list()
for (comparison_name in names(inter_tissue_results)) {
  comparison_results <- inter_tissue_results[[comparison_name]]
  
  # Add the data frame to the list
  df_list[[paste0(comparison_name, "_", chromosome_name)]] <- comparison_results
}

inter_df <- do.call(rbind, df_list)
fwrite(inter_df, "Inter-tissue_differential_TAD_bounds.txt", sep="\t", quote=F, row.names=F, col.names=T)
```

#Old, scratch from chat w/ GitHub CoPilot
```{r}
# Extract the sample number IDs from the paths
AFR_ids <- basename(AFR_paths)
EUR_ids <- basename(EUR_paths)

# Get the current names in results
names <- names(results)

# Split the names at the underscore
split_names <- strsplit(names, "_")

# Replace the indices with the sample number IDs
new_names <- mapply(function(name, afr, eur) {
  parts <- strsplit(name, "_")[[1]]
  parts[c(2, 4)] <- c(afr, eur)
  paste(parts, collapse = "_")
}, names, rep(AFR_ids, each=length(EUR_ids)), rep(EUR_ids, times=length(AFR_ids)))

names(results) <- new_names

saveRDS(results, "TADCompare_results.rds")














# Get the list of all AFR and EUR samples
AFR_samples <- list(SSAMPLE_09, SSAMPLE_10, SSAMPLE_11, SSAMPLE_12, SSAMPLE_13, SSAMPLE_14, SSAMPLE_15)
EUR_samples <- list(SSAMPLE_16, SSAMPLE_17, SSAMPLE_18, SSAMPLE_19, SSAMPLE_20, SSAMPLE_21, SSAMPLE_22, SSAMPLE_05)

# Initialize an empty list to store the results
results <- list()

test <- TADCompare(SSAMPLE_09[[22]][1], SSAMPLE_16[[22]][1], resolution=50000)


TAD_comparer <- function(path1, path2, res){
  bed <- read.table(paste0(path1, ".", res, "_abs.bed"))
  bed[,1] <- paste0("chr", bed[,1])
  mat <- read.table(paste0(path1, ".", res, ".matrix"))
  dat1 <- HiCcompare::hicpro2bedpe(mat, bed)
  dat1 <- dat1$cis
  dat1 <- dat1[-grep("Un|alt|random|X|Y|M", names(dat1), perl=T)]
  
  bed <- read.table(paste0(path2, ".", res, "_abs.bed"))
  bed[,1] <- paste0("chr", bed[,1])
  mat <- read.table(paste0(path2, ".", res, ".matrix"))
  dat2 <- HiCcompare::hicpro2bedpe(mat, bed)
  dat2 <- dat2$cis
  dat2 <- dat2[-grep("Un|alt|random|X|Y|M", names(dat2), perl=T)]
  
  sparse_tads = lapply(1:length(dat1), function(z) {
    x <- dat1[[z]]
    y <- dat2[[z]]
    
    #Pull out chromosome
    chr <- x[, 1][1]
    #Subset to make three column matrix
    x <- x[, c(2, 5, 7)]
    y <- y[, c(2, 5, 7)]

    #Run SpectralTAD
    comp <- TADCompare(x, y, resolution = 50000)

    return(list("comp"=comp, "chr"=chr))
  })
  diff_res <- lapply(sparse_tads, function(x) x[["comp"]])
  chr <- lapply(sparse_tads, function(x) x[["chr"]])
  names(diff_res) <- chr
  return(diff_res)
}

test <- TAD_comparer("<path>/FULL_SAMPLES/SAMPLE_09/raw/SAMPLE_09", "<path>/FULL_SAMPLES/SAMPLE_16/raw/SAMPLE_16", 50000)




# Function to read and format data
read_format_data <- function(path, res) {
  bed <- read.table(paste0(path, ".", res, "_abs.bed"))
  bed[,1] <- paste0("chr", bed[,1])
  mat <- read.table(paste0(path, ".", res, ".matrix"))
  dat <- HiCcompare::hicpro2bedpe(mat, bed)
  dat <- dat$cis
  dat <- dat[-grep("Un|alt|random|X|Y|M", names(dat), perl=T)]
  return(dat)
}

# Function to make pairwise comparisons
compare_pairwise <- function(dat1, dat2) {
  sparse_tads = lapply(1:length(dat1), function(z) {
    x <- dat1[[z]]
    y <- dat2[[z]]
    
    # Pull out chromosome
    chr <- x[, 1][1]
    # Subset to make three column matrix
    x <- x[, c(2, 5, 7)]
    y <- y[, c(2, 5, 7)]
    
    # Run SpectralTAD
    comp <- TADCompare(x, y, resolution = 50000)
    return(list("comp" = comp, "chr" = chr))
  })
  
  diff_res <- lapply(sparse_tads, function(x) x[["comp"]])
  chr <- lapply(sparse_tads, function(x) x[["chr"]])
  names(diff_res) <- chr
  return(diff_res)
}






# Loop over all AFR and EUR samples
for (i in seq_along(AFR_samples)) {
  for (j in seq_along(EUR_samples)) {
    pair_results <- lapply(1:22, function(k) {
      cat("i:", i, "j:", j, "k:", k, "\n")
      print(head(AFR_samples[[i]][[k]][1]))
      #cat("AFR_samples[[i]][[k]][1]:", AFR_samples[[i]][[k]][1], "\n")
      #cat("EUR_samples[[j]][[k]][1]:", EUR_samples[[j]][[k]][1], "\n")
      result <- TADCompare(AFR_samples[[i]][[k]][1], EUR_samples[[j]][[k]][1], resolution=50000)
      result$chromosome <- paste0("chr", k)
      result$pair <- paste0("AFR", i, "_EUR", j)
      result
    })
    results[[paste0("AFR", i, "_EUR", j)]] <- do.call(rbind, pair_results)
  }
}

# Combine all results into a single data frame
df <- do.call(rbind, results)

# Write the data frame to a file
fwrite(df, "TADCompare_results.txt", sep="\t", quote=F, row.names=F, col.names=T)
```

## Differential Loop Call Functions
```{r}
#HiCCompare functions
dat_reader <- function(path, res) {
  bed <- read.table(paste0(path, "sample_", res, "_abs.bed"))
  bed[,1] <- paste0("chr", bed[,1])
  mat <- read.table(paste0(path, "sample_", res, ".matrix"))
  dat <- HiCcompare::hicpro2bedpe(mat, bed)
  dat <- dat$cis
  dat <- dat[-grep("Un|alt|random|X|Y|M", names(dat), perl=T)]
  return(dat)
}

dat_normalizer <- function(dat1, dat2){
  hic.list <- mapply(create.hic.table, dat1, dat2, SIMPLIFY=FALSE, scale=FALSE)
  hic.list <- total_sum(hic.list)
  hic.table <- hic_loess(hic.list, Plot=TRUE, parallel=TRUE)
  return(hic.table)
}

dat_comparer <- function(hic_table, writepath, A_min=NA){
  hic.table <- hic_compare(hic_table, A.min=A_min, Plot=TRUE, parallel=TRUE)
  objout <- bind_rows(hic.table)
  objout <- objout[order(objout$p.adj),]
  objout <- objout[which(objout$p.adj<=0.05),]
  fwrite(objout, writepath, sep="\t", row.names=F, col.names=T, quote=F)
  return(hic.table)
}

dat_analyzer <- function(datapath1, datapath2, res, outpath){
  dat1 <- dat_reader(paste0(datapath1, "/", res, "/"), res)
  dat2 <- dat_reader(paste0(datapath2, "/", res, "/"), res)
  dat1v2 <- dat_normalizer(dat1, dat2)
  dat_res <- dat_comparer(dat1v2, paste0(outpath, "_", res, ".txt"))
  return(dat_res)
}
```

## Differential loop calls, 5kb!
```{r}
options(scipen=999)
dat_analyzer("<path>/FULL_SAMPLES/merged_data_black_men/hic_results/matrix/sample/raw", "<path>/FULL_SAMPLES/merged_data_black_women/hic_results/matrix/sample/raw", 5000, "<path>/FULL_SAMPLES/merged_bm_vs_bw")

dat_analyzer("<path>/FULL_SAMPLES/merged_data_white_men/hic_results/matrix/sample/raw", "<path>/FULL_SAMPLES/merged_data_white_women/hic_results/matrix/sample/raw", 5000, "<path>/FULL_SAMPLES/merged_wm_vs_ww")

dat_analyzer("<path>/FULL_SAMPLES/merged_data_black_men/hic_results/matrix/sample/raw", "<path>/FULL_SAMPLES/merged_data_white_men/hic_results/matrix/sample/raw", 5000, "<path>/FULL_SAMPLES/merged_bm_vs_wm")

dat_analyzer("<path>/FULL_SAMPLES/merged_data_black_women/hic_results/matrix/sample/raw", "<path>/FULL_SAMPLES/merged_data_white_women/hic_results/matrix/sample/raw", 5000, "<path>/FULL_SAMPLES/merged_bw_vs_ww")

dat_analyzer("<path>/FULL_SAMPLES/merged_data_men/hic_results/matrix/sample/raw", "<path>/FULL_SAMPLES/merged_data_women/hic_results/matrix/sample/raw", 5000, "<path>/FULL_SAMPLES/merged_m_vs_w")

dat_analyzer("<path>/FULL_SAMPLES/merged_data_white/hic_results/matrix/sample/raw", "<path>/FULL_SAMPLES/merged_data_black/hic_results/matrix/sample/raw", 5000, "<path>/FULL_SAMPLES/merged_w_vs_b")
```

## Multi Hi-C Compare, looking at individual-level data instead (may be more reliable!)
```{r}

#MultiHiCCompare functions
dat_reader_mult <- function(path, res) {
  bed <- read.table(paste0(path, ".", res, "_abs.bed"))
  bed[,1] <- paste0("chr", bed[,1])
  mat <- read.table(paste0(path, ".", res, ".matrix"))
  dat <- HiCcompare::hicpro2bedpe(mat, bed)
  dat <- dat$cis
  dat <- dat[-grep("Un|alt|random|X|Y|M", names(dat), perl=T)]
  final <- bind_rows(dat) %>% select(., chr1, start1, start2, IF)
  colnames(final) <- c("chr", "region1", "region2", "IF")
  return(final)
}

dat_normalizer_mult <- function(dat1, dat2, groups){
  hicexp <- make_hicexp(dat1, dat2, groups=groups, remove.regions=hg38_cyto)
  hicexp <- cyclic_loess(hicexp, verbose=FALSE, parallel=TRUE)
  print(MD_hicexp(hicexp))
}

dat_reader_mult_chr <- function(path, res, chr) {
  bed <- read.table(paste0(path, ".", res, "_", chr, "_abs.bed"))
  bed[,1] <- paste0("chr", bed[,1])
  mat <- read.table(paste0(path, ".", res, "_", chr, ".matrix"))
  dat <- HiCcompare::hicpro2bedpe(mat, bed)
  dat <- dat$cis
  final <- bind_rows(dat) %>% select(., chr1, start1, start2, IF)
  colnames(final) <- c("chr", "region1", "region2", "IF")
  
  na_inf_indices <- which(is.na(final$IF) | final$IF== Inf)
  print(length(na_inf_indices))
  
  return(final)
}
```

### MultiHiCCompare testing
```{r}
options(scipen=999)

#Troubleshooting:
SSAMPLE_16 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_16/raw/SAMPLE_16", 5000, "X")
SSAMPLE_17 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_17/raw/SAMPLE_17", 5000, "X")
SSAMPLE_18 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_18/raw/SAMPLE_18", 5000, "X")
SSAMPLE_19 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_19/raw/SAMPLE_19", 5000, "X")
SSAMPLE_20 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_20/raw/SAMPLE_20", 5000, "X")
SSAMPLE_21 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_21/raw/SAMPLE_21", 5000, "X")
SSAMPLE_22 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_22/raw/SAMPLE_22", 5000, "X")
SSAMPLE_05 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_05/raw/SAMPLE_05", 5000, "X")

SSAMPLE_09 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_09/raw/SAMPLE_09", 5000, "X")
SSAMPLE_10 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_10/raw/SAMPLE_10", 5000, "X")
SSAMPLE_11 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_11/raw/SAMPLE_11", 5000, "X")
SSAMPLE_12 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_12/raw/SAMPLE_12", 5000, "X")
SSAMPLE_13 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_13/raw/SAMPLE_13", 5000, "X")
SSAMPLE_14 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_14/raw/SAMPLE_14", 5000, "X")
SSAMPLE_15 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_15/raw/SAMPLE_15", 5000, "X")

#ChrX for females
myexp <- make_hicexp(SSAMPLE_09, SSAMPLE_10, SSAMPLE_11, SSAMPLE_16, SSAMPLE_17, SSAMPLE_18, SSAMPLE_05, groups=c(0,0,0,1,1,1,1), remove.regions=hg38_cyto)
myexp <- multiHiCcompare::fastlo(myexp, parallel=TRUE, verbose=TRUE)
myexp <- hic_exactTest(myexp, p.method="fdr", parallel=TRUE)
final <- results(myexp)
final <- final[order(final$p.adj),]
final[which(final$p.adj<=0.05),]

#all_bins <- unique(rbind(SSAMPLE_16, SSAMPLE_17, SSAMPLE_18, SSAMPLE_19, SSAMPLE_20, SSAMPLE_21, SSAMPLE_22, SSAMPLE_05, SSAMPLE_09, SSAMPLE_10, SSAMPLE_11, SSAMPLE_12, SSAMPLE_13, SSAMPLE_14, SSAMPLE_15)[,c("chr", "region1", "region2")])

# Define a function to add missing bins
add_missing_bins <- function(dat) {
  # Find the bins that are missing in this dataset
  missing_bins <- setdiff(all_bins, dat[, c("chr", "region1", "region2")])

  # Create a data frame with zero counts for the missing bins
  missing_data <- data.frame(missing_bins, IF = 1E-10)

  # Add the missing data to the dataset
  dat <- rbind(dat, missing_data)

  return(dat)
}

#SSAMPLE_16 <- add_missing_bins(SSAMPLE_16)
#SSAMPLE_17 <- add_missing_bins(SSAMPLE_17)
#SSAMPLE_18 <- add_missing_bins(SSAMPLE_18)
#SSAMPLE_19 <- add_missing_bins(SSAMPLE_19)
#SSAMPLE_20 <- add_missing_bins(SSAMPLE_20)
#SSAMPLE_21 <- add_missing_bins(SSAMPLE_21)
#SSAMPLE_22 <- add_missing_bins(SSAMPLE_22)
#SSAMPLE_05 <- add_missing_bins(SSAMPLE_05)
#SSAMPLE_09 <- add_missing_bins(SSAMPLE_09)
#SSAMPLE_10 <- add_missing_bins(SSAMPLE_10)
#SSAMPLE_11 <- add_missing_bins(SSAMPLE_11)
#SSAMPLE_12 <- add_missing_bins(SSAMPLE_12)
#SSAMPLE_13 <- add_missing_bins(SSAMPLE_13)
#SSAMPLE_14 <- add_missing_bins(SSAMPLE_14)
#SSAMPLE_15 <- add_missing_bins(SSAMPLE_15)

myexp <- make_hicexp(SSAMPLE_09, SSAMPLE_10, SSAMPLE_11, SSAMPLE_12, SSAMPLE_13, SSAMPLE_14, SSAMPLE_15, SSAMPLE_16, SSAMPLE_17, SSAMPLE_18, SSAMPLE_19, SSAMPLE_20, SSAMPLE_21, SSAMPLE_22, SSAMPLE_05, groups=c(0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1), remove.regions=hg38_cyto)
myexp <- multiHiCcompare::fastlo(myexp, parallel=TRUE, verbose=TRUE)
myexp <- hic_exactTest(myexp, p.method="fdr", parallel=TRUE)

#Get chr18 results at 1kb! AA vs. EUR!
myexp2 <- make_hicexp(SSAMPLE_09, SSAMPLE_10, SSAMPLE_11, SSAMPLE_12, SSAMPLE_13, SSAMPLE_15, SSAMPLE_16, SSAMPLE_17, SSAMPLE_18, SSAMPLE_19, SSAMPLE_20, SSAMPLE_21, SSAMPLE_22, SSAMPLE_05, groups=c(0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1, 1), remove.regions=hg38_cyto)
myexp2 <- multiHiCcompare::fastlo(myexp2, parallel=TRUE, verbose=TRUE)
myexp2 <- hic_exactTest(myexp2, p.method="fdr", parallel=TRUE)
myexp2_res <- results(myexp2) %>% filter(., p.adj<=0.1)
fwrite(myexp2_res, "AA_vs_EUR_1kb_chr18_result.txt", sep="\t", row.names=F, col.names=T, quote=F)

#Get chr18 results at 1kb, M v W:
myexp3 <- make_hicexp(SSAMPLE_12, SSAMPLE_13, SSAMPLE_15, SSAMPLE_19, SSAMPLE_20, SSAMPLE_21, SSAMPLE_22, SSAMPLE_09, SSAMPLE_10, SSAMPLE_11, SSAMPLE_05, SSAMPLE_16, SSAMPLE_17, SSAMPLE_18, groups=c(0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1), remove.regions=hg38_cyto)
myexp3 <- multiHiCcompare::fastlo(myexp3, parallel=TRUE, verbose=TRUE)
myexp3 <- hic_exactTest(myexp3, p.method="fdr", parallel=TRUE)
myexp3_res <- results(myexp3) %>% filter(., p.adj<=0.1)
fwrite(myexp3_res, "M_vs_W_1kb_chr18_result.txt", sep="\t", row.names=F, col.names=T, quote=F)

#Repeat the above, for chr5, re-including the dropped sample:
SSAMPLE_16 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_16/raw/SAMPLE_16/SAMPLE_16", 1000, 5)
SSAMPLE_17 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_17/raw/SAMPLE_17/SAMPLE_17", 1000, 5)
SSAMPLE_18 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_18/raw/SAMPLE_18/SAMPLE_18", 1000, 5)
SSAMPLE_19 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_19/raw/SAMPLE_19/SAMPLE_19", 1000, 5)
SSAMPLE_20 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_20/raw/SAMPLE_20/SAMPLE_20", 1000, 5)
SSAMPLE_21 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_21/raw/SAMPLE_21/SAMPLE_21", 1000, 5)
SSAMPLE_22 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_22/raw/SAMPLE_22/SAMPLE_22", 1000, 5)
SSAMPLE_05 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_05/raw/SAMPLE_05/SAMPLE_05", 1000, 5)

SSAMPLE_09 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_09/raw/SAMPLE_09/SAMPLE_09", 1000, 5)
SSAMPLE_10 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_10/raw/SAMPLE_10/SAMPLE_10", 1000, 5)
SSAMPLE_11 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_11/raw/SAMPLE_11/SAMPLE_11", 1000, 5)
SSAMPLE_12 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_12/raw/SAMPLE_12/SAMPLE_12", 1000, 5)
SSAMPLE_13 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_13/raw/SAMPLE_13/SAMPLE_13", 1000, 5)
SSAMPLE_14 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_14/raw/SAMPLE_14/SAMPLE_14", 1000, 5)
SSAMPLE_15 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_15/raw/SAMPLE_15/SAMPLE_15", 1000, 5)

#Get chr5 results at 1kb, M v W:
myexp4 <- make_hicexp(SSAMPLE_12, SSAMPLE_13, SSAMPLE_14, SSAMPLE_15, SSAMPLE_19, SSAMPLE_20, SSAMPLE_21, SSAMPLE_22, SSAMPLE_09, SSAMPLE_10, SSAMPLE_11, SSAMPLE_05, SSAMPLE_16, SSAMPLE_17, SSAMPLE_18, groups=c(0, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 1, 1, 1, 1), remove.regions=hg38_cyto)
myexp4 <- multiHiCcompare::fastlo(myexp4, parallel=TRUE, verbose=TRUE)
myexp4 <- hic_exactTest(myexp4, p.method="fdr", parallel=TRUE)
myexp4_res <- results(myexp4) %>% filter(., p.adj<=0.1)
fwrite(myexp4_res, "M_vs_W_1kb_chr5_result.txt", sep="\t", row.names=F, col.names=T, quote=F)

#Walk through this step-by-step based on https://github.com/dozmorovlab/multiHiCcompare/blob/dcfe4aaa8eaef45e203f3d7f806232bb613d2c9b/R/glm.R#L34
table_list <- split(hic_table(myexp), hic_table(myexp)$chr)
table_list <- lapply(table_list, .get_dist_tables, max.pool=0.7)
table_list <- do.call(c, table_list)
dge_list <- lapply(table_list, .hictable2DGEList, covariates=meta(myexp))

#Estimate dispersion, this is where I think things go awry:
if ( length(meta(myexp)$group) == 2) { # IF no replicates use edgeR's recommended method to estimate dispersion
    dge_list <- smartApply(parallel, dge_list, edgeR::estimateGLMCommonDisp,
                                                          method="deviance", robust=TRUE, 
                                                          subset=NULL)
  } else { # If replicates for condition then use standard way for estimating dispersion
    dge_list <- smartApply(parallel, dge_list, edgeR::estimateDisp, 
                           design=model.matrix(~meta(hicexp)$group))

  }

inter_data <- multiHiCcompare::interactions(myexp)


#Validate nf pipeline results:
SSAMPLE_09 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_09/raw/SAMPLE_09/SAMPLE_09", 5000, 20)
SSAMPLE_10 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_10/raw/SAMPLE_10/SAMPLE_10", 5000, 20)
SSAMPLE_11 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_11/raw/SAMPLE_11/SAMPLE_11", 5000, 20)
SSAMPLE_12 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_12/raw/SAMPLE_12/SAMPLE_12", 5000, 20)
SSAMPLE_13 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_13/raw/SAMPLE_13/SAMPLE_13", 5000, 20)
SSAMPLE_14 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_14/raw/SAMPLE_14/SAMPLE_14", 5000, 20)
SSAMPLE_15 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_15/raw/SAMPLE_15", 5000, 20)

SSAMPLE_16 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_16/raw/SAMPLE_16/SAMPLE_16", 5000, 20)
SSAMPLE_17 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_17/raw/SAMPLE_17/SAMPLE_17", 5000, 20)
SSAMPLE_18 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_18/raw/SAMPLE_18/SAMPLE_18", 5000, 20)
SSAMPLE_19 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_19/raw/SAMPLE_19/SAMPLE_19", 5000, 20)
SSAMPLE_20 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_20/raw/SAMPLE_20/SAMPLE_20", 5000, 20)
SSAMPLE_21 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_21/raw/SAMPLE_21/SAMPLE_21", 5000, 20)
SSAMPLE_22 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_22/raw/SAMPLE_22/SAMPLE_22", 5000, 20)
SSAMPLE_05 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_05/raw/SAMPLE_05/SAMPLE_05", 5000, 20)

test <- make_hicexp(SSAMPLE_09, SSAMPLE_10, SSAMPLE_11, SSAMPLE_12, SSAMPLE_13, SSAMPLE_14, SSAMPLE_15, SSAMPLE_16, SSAMPLE_17, SSAMPLE_18, SSAMPLE_19, SSAMPLE_20, SSAMPLE_21, SSAMPLE_22, SSAMPLE_05, groups=c(0,0,0,0,0,0,0,1,1,1,1,1,1,1,1), remove.regions=hg38_cyto)
test <- fastlo(test, parallel=TRUE, verbose=TRUE)
test <- hic_exactTest(test, p.method="fdr", parallel = TRUE)
final <- results(test)
final <- filter(final, p.adj<=0.1)
print(final)



finalall <- NULL
for(CHR in c(1:22)) {
  SSAMPLE_19 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_19/raw/SAMPLE_19", 5000, CHR)
  SSAMPLE_20 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_20/raw/SAMPLE_20", 5000, CHR)
  SSAMPLE_12 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_12/raw/SAMPLE_12", 5000, CHR)
  SSAMPLE_13 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_13/raw/SAMPLE_13", 5000, CHR)
  
  myexp <- make_hicexp(SSAMPLE_19, SSAMPLE_20, SSAMPLE_12, SSAMPLE_13, groups=c(0, 0, 1, 1), remove.regions=hg38_cyto)
  myexp <- fastlo(myexp, parallel=TRUE, verbose=TRUE)
  print(MD_hicexp(myexp))
  myexp <- hic_exactTest(myexp, p.method="fdr", parallel=TRUE)
  final <- results(myexp)
  final <- filter(final, p.adj<=0.05)
  finalall <- rbind(finalall, final)
}

SSAMPLE_09
SSAMPLE_19.5000.22 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_19/raw/SAMPLE_19", 5000, 22)
SSAMPLE_20.5000.22 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_20/raw/SAMPLE_20", 5000, 22)
SSAMPLE_12.5000.22 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_12/raw/SAMPLE_12", 5000, 22)
SSAMPLE_13.5000.22 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_13/raw/SAMPLE_13", 5000, 22)

myexp <- make_hicexp(SSAMPLE_19.5000.22, SSAMPLE_20.5000.22, SSAMPLE_12.5000.22, SSAMPLE_13.5000.22, groups=c(0, 0, 1, 1), remove.regions=hg38_cyto)
myexp <- fastlo(myexp, parallel=TRUE, verbose=TRUE)
MD_hicexp(myexp)
myexp <- hic_exactTest(myexp, p.method="fdr", parallel=TRUE)
final <- results(myexp)
final <- filter(final, p.adj<=0.05)

allres <- NULL
allres <- rbind(allres, final)

SSAMPLE_19.5000.18 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_19/raw/SAMPLE_19", 5000, 18)
SSAMPLE_20.5000.18 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_20/raw/SAMPLE_20", 5000, 18)
SSAMPLE_12.5000.18 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_12/raw/SAMPLE_12", 5000, 18)
SSAMPLE_13.5000.18 <- dat_reader_mult_chr("<path>/FULL_SAMPLES/SAMPLE_13/raw/SAMPLE_13", 5000, 18)

myexp <- make_hicexp(SSAMPLE_19.5000.18, SSAMPLE_20.5000.18, SSAMPLE_12.5000.18, SSAMPLE_13.5000.18, groups=c(0, 0, 1, 1), remove.regions=hg38_cyto)
myexp <- fastlo(myexp, parallel=TRUE, verbose=TRUE)
MD_hicexp(myexp)
myexp <- hic_exactTest(myexp, p.method="fdr", parallel=TRUE)
final18 <- results(myexp)
final18 <- filter(final18, p.adj<=0.05)

allres <- rbind(allres, final18)

#Caucasian females
SSAMPLE_05 <- dat_reader_mult("<path>/FULL_SAMPLES/SAMPLE_05/raw/SAMPLE_05", 500000)
SSAMPLE_16 <- dat_reader_mult("<path>/FULL_SAMPLES/SAMPLE_16/raw/SAMPLE_16", 500000)
SSAMPLE_17 <- dat_reader_mult("<path>/FULL_SAMPLES/SAMPLE_17/raw/SAMPLE_17", 500000)
SSAMPLE_18 <- dat_reader_mult("<path>/FULL_SAMPLES/SAMPLE_18/raw/SAMPLE_18", 500000)

#AA females
SSAMPLE_09 <- dat_reader_mult("<path>/FULL_SAMPLES/SAMPLE_09/raw/SAMPLE_09", 500000)
SSAMPLE_10 <- dat_reader_mult("<path>/FULL_SAMPLES/SAMPLE_10/raw/SAMPLE_10", 500000)
SSAMPLE_11 <- dat_reader_mult("<path>/FULL_SAMPLES/SAMPLE_11/raw/SAMPLE_11", 500000)

mygroups="0, 0, 0, 0, 1, 1, 1"
mg <- as.numeric(strsplit(mygroups, split=", ", fixed=TRUE)[[1]])

hicexp_test <- make_hicexp(SSAMPLE_05, SSAMPLE_16, SSAMPLE_17, SSAMPLE_18, SSAMPLE_09, SSAMPLE_10, SSAMPLE_11, groups=mg, remove.regions=hg38_cyto)
hicexp_test <- fastlo(hicexp_test, parallel=TRUE)
MD_hicexp(hicexp_test)

hicexp_test <- hic_exactTest(hicexp_test, p.method = 'fdr', parallel = TRUE)

# plot results
MD_composite(hicexp_test)
results(hicexp_test)
```
